---
title: "Get started with precrec"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Get started with precrec}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`precrec` calculates and plots ROC and precision-recall curves for binary
classifiers. It is built for the case where the two curves disagree: on an
imbalanced dataset a ROC curve can look excellent while the precision-recall
curve shows the classifier is not usable.

This page is the five-minute tour. Everything else lives on the
[package website](https://evalclass.github.io/precrec/).

## The one function you need

`evalmod()` takes scores and labels and returns an object the plotting and
summary functions understand.

```{r}
library(precrec)

# 10 positives and 10 negatives, shipped with the package
data(P10N10)

curves <- evalmod(scores = P10N10$scores, labels = P10N10$labels)
```

`scores` are the classifier's predictions - any numeric value, higher meaning
more likely positive. `labels` are the observed classes. Neither has to be
sorted, and the scores do not have to be probabilities.

## Look at it

```{r, fig.width = 7, fig.height = 4}
plot(curves)
```

`autoplot()` draws the same thing with `ggplot2`, which is the one to use if
you want to restyle the result.

```{r, fig.width = 7, fig.height = 4}
library(ggplot2)

autoplot(curves)
```

## Get the numbers out

`auc()` returns the areas under both curves.

```{r}
knitr::kable(auc(curves))
```

`as.data.frame()` returns the curve points themselves, ready for any other
tool.

```{r}
head(as.data.frame(curves))
```

## Where to go next

The website has three sets of short pages.

| Section | What is in it |
|---|---|
| [How-to](https://evalclass.github.io/precrec/articles/) | One page per task: several models, several test sets, cross-validation, more than two classes, large datasets |
| [Measures](https://evalclass.github.io/precrec/articles/measures-overview.html) | What each of the 26 available measures means and when it misleads |
| [Plots](https://evalclass.github.io/precrec/articles/plots-overview.html) | Every plot the package draws, and how to change it |

## Citation

*Precrec: fast and accurate precision-recall and ROC curve calculations in R*

Takaya Saito; Marc Rehmsmeier

Bioinformatics 2017; 33 (1): 145-147.
doi: [10.1093/bioinformatics/btw570](https://doi.org/10.1093/bioinformatics/btw570)
