---
title: "Evaluate cross-validation folds"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Cross-validation folds are just several test sets, so everything on
[averaging over test sets](howto-multiple-test-sets.html) applies. The only
extra step is getting your data out of a data frame.

```{r}
library(precrec)
library(ggplot2)
```

## The usual shape of the data

`M2N50F5` holds two models scored on 5 folds - one column per model, one
label column and one fold column.

```{r}
data(M2N50F5)

knitr::kable(head(M2N50F5))
```

## Straight into evalmod

`evalmod()` and `mmdata()` both take the fold columns directly. This is the
short way.

```{r}
curves <- evalmod(
  nfold_df = M2N50F5, score_cols = c(1, 2),
  lab_col = 3, fold_col = 4,
  modnames = c("m1", "m2"), dsids = 1:5
)

autoplot(curves)
```

Column names work as well as positions.

```{r}
curves2 <- evalmod(
  nfold_df = M2N50F5,
  score_cols = c("score1", "score2"),
  lab_col = "label", fold_col = "fold",
  modnames = c("m1", "m2"), dsids = 1:5
)
```

## Converting first

`format_nfold()` does the conversion on its own if you want the lists for
something else.

```{r}
folds <- format_nfold(
  nfold_df = M2N50F5, score_cols = c(1, 2),
  lab_col = 3, fold_col = 4
)

curves3 <- evalmod(
  scores = folds$scores, labels = folds$labels,
  modnames = rep(c("m1", "m2"), each = 5),
  dsids = rep(1:5, 2)
)
```

## Reading the result

Each model gets one averaged curve with a confidence band across the folds.

```{r}
knitr::kable(auc_ci(curves))
```

## Next

- [Average over several test sets](howto-multiple-test-sets.html)
- [Get the numbers out](howto-results-as-data.html)
