---
title: "Evaluate more than two classes"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

A dataset with more than two classes is evaluated by one-vs-rest
decomposition: each class becomes its own binary problem - that class against
all the others - so the accurate precision-recall calculations apply to it
unchanged.

```{r}
library(precrec)
library(ggplot2)
```

## The input

Pass a matrix with one score column per class, together with the class
labels.

```{r}
# A 3-class dataset with one score column per class
data(C3N150)

mdat <- mmdata(C3N150$scores, C3N150$labels)
mdat
```

The decomposition is detected from the input. The classes are carried on the
model axis, so everything downstream treats them the way it treats several
models on one test set.

Add `multiclass = "ovr"` to ask for the decomposition explicitly. That is
worth doing when the score columns are not named after the classes - they are
then taken in the order of the class names.

## One curve per class

```{r}
curves <- evalmod(mdat)

autoplot(curves)
```

## Per-class and macro-averaged areas

`auc()` reports one row per class and curve type, then the macro-average of
the per-class scores.

```{r}
knitr::kable(auc(curves))
```

`macro = FALSE` leaves the average out and reports the classes alone.

The average weights every class equally by default, so a class with ten
observations counts as much as one with a thousand. `macro_weight =
"prevalence"` weights each class by its size instead.

```{r}
knitr::kable(auc(curves, macro_weight = "prevalence"))
```

Use the uniform average when every class matters equally - usually the case
when the rare classes are the interesting ones - and the weighted average
when you want the average case. Other packages call these `roc_aunu` and
`roc_aunp`.

## One thing to watch

Each one-vs-rest split has its own class balance, so the baseline of a
precision-recall curve differs from class to class. The plots leave the
baseline out for that reason - which matters here more than anywhere else,
because a precision-recall curve is read against its baseline.

## Next

- [Balanced and imbalanced data](howto-imbalanced-data.html)
- [AUC and other curve summaries](metrics-auc.html)
