---
title: "Compare several models"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Several models tested on the *same* data give one curve each, drawn together
on one plot.

```{r}
library(precrec)
library(ggplot2)

samps <- create_sim_samples(1, 100, 100, "all")

mdat <- mmdata(samps[["scores"]], samps[["labels"]],
  modnames = samps[["modnames"]]
)
```

The five models here are the five simulated quality levels, from random to
perfect.

## Calculate and plot

`evalmod()` notices there is more than one model and labels the curves
accordingly.

```{r, fig.height = 4}
curves <- evalmod(mdat)

autoplot(curves)
```

## Compare the areas

```{r}
knitr::kable(auc(curves))
```

The gap between the two curve types is the point of the package: ROC areas
stay high for models the precision-recall areas show to be weak. See
[balanced and imbalanced data](howto-imbalanced-data.html).

## One curve type at a time

```{r, fig.height = 4}
autoplot(curves, "PRC")
```

## Naming the models

Without `modnames`, models are named `m1`, `m2` and so on. Pass your own
names to `mmdata()` or straight to `evalmod()`.

```{r}
curves2 <- evalmod(
  scores = samps[["scores"]], labels = samps[["labels"]],
  modnames = c("random", "poor", "good", "excellent", "perfect")
)
```

## Next

- [Average over several test sets](howto-multiple-test-sets.html)
- [Get the numbers out](howto-results-as-data.html)
