---
title: "Get the numbers out"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Every `precrec` object converts to a data frame, and every summary is a data
frame too. Nothing is locked inside the plotting code.

```{r}
library(precrec)

curves <- evalmod(scores = P10N10$scores, labels = P10N10$labels)
```

## What is in the object

`print()` summarizes the input and the results.

```{r}
curves
```

## The curve points

```{r}
df <- as.data.frame(curves)
head(df)
```

One row per supporting point, with the curve type and the model in their own
columns - the shape `ggplot2` and `dplyr` expect. `as.data.table()` returns
the same thing as a `data.table` when that package is installed.

## The summaries

| Function | Returns |
|---|---|
| `auc()` | Area under each curve |
| `pauc()` | Partial area, after `part()` |
| `auc_ci()` | Confidence interval of the area, over several test sets |
| `prbe()` | Precision-recall break-even point |
| `prob_metrics()` | Brier score, RMSE and log loss |

```{r}
knitr::kable(auc(curves))
```

Each returns a plain data frame, so subsetting is ordinary R.

```{r}
aucs <- auc(curves)
knitr::kable(subset(aucs, curvetypes == "PRC"))
```

## Basic measures per cutoff

`mode = "basic"` gives the per-cutoff measures rather than the curves.

```{r}
points <- evalmod(scores = P10N10$scores, labels = P10N10$labels,
  mode = "basic"
)

head(as.data.frame(points))
```

The `x` column is the normalized rank and `y` the value of the measure named
in `type`.

## Feeding ggplot2 directly

`fortify()` is the `ggplot2` hook, so a `precrec` object can go straight into
`ggplot()`. See [customizing plots](plots-customizing.html).

## Next

- [Measures overview](measures-overview.html)
- [Plots overview](plots-overview.html)
