---
title: "Average over several test sets"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

One model tested on several datasets gives several curves. `precrec` averages
them and draws a confidence band around the average.

```{r}
library(precrec)
library(ggplot2)

# 10 test sets for one model
samps <- create_sim_samples(10, 100, 100, "good_er")

mdat <- mmdata(samps[["scores"]], samps[["labels"]], dsids = samps[["dsids"]])
```

`dsids` is what marks these as several test sets rather than several models.

## The average curve

```{r}
curves <- evalmod(mdat)

autoplot(curves, "PRC")
```

The shaded band is the point-wise 95% confidence interval of the average. It
appears on its own as soon as a model has more than one test set.

## The individual curves

Ask for them with `raw_curves = TRUE`.

```{r}
raw <- evalmod(mdat, raw_curves = TRUE)

autoplot(raw, "PRC", show_cb = FALSE)
```

## Averaged areas

`auc()` reports the mean area over the test sets.

```{r}
knitr::kable(auc(curves))
```

`auc_ci()` puts a confidence interval on it.

```{r}
knitr::kable(auc_ci(curves))
```

## Both at once

Several models *and* several test sets is the same call with both
identifiers set. Each model gets one average curve.

```{r}
samps2 <- create_sim_samples(10, 100, 100, c("poor_er", "good_er"))

mdat2 <- mmdata(samps2[["scores"]], samps2[["labels"]],
  modnames = samps2[["modnames"]], dsids = samps2[["dsids"]]
)

autoplot(evalmod(mdat2), "PRC")
```

## Next

- [Confidence bands](plots-confidence-bands.html) - how the band is built and
  how to change it
- [Evaluate cross-validation folds](howto-cross-validation.html)
