---
title: "Choose an operating point"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Almost everything else here is threshold-free. A curve is every cutoff at
once and an AUC summarizes all of them, which is what makes them fair ways
to compare models. Shipping one means giving that up: something has to turn
a score into a decision, and that something is a threshold.

`best_cutoff()` picks it.

```{r}
library(precrec)

set.seed(1)
scores <- c(rnorm(20, 1.2), rnorm(380, 0))
labels <- rep(c(1, 0), c(20, 380))
mdat <- mmdata(scores, labels)

best_cutoff(mdat)
```

## What comes back is a row, not a number

The threshold alone would not tell you whether to accept it. What comes back
is a row of [`metric_table()`](howto-results-as-data.html), so every other
metric is there to be read at the same cutoff:

```{r}
picked <- best_cutoff(mdat, metric = "mcc")

knitr::kable(
  picked[, c(
    "metric", "value", "rank", "score",
    "sensitivity", "specificity", "precision"
  )],
  digits = 3
)
```

`rank` is how many of the 400 instances that threshold calls positive, and
`score` is the cutoff itself - an observed score, with the rule being
`score >= ` it. Reading the precision beside the criterion is the point of
returning the whole row: a cutoff can optimize what you asked for and still
be one you would not deploy.

## The criteria come in two groups

```{r}
criteria <- c("youden", "topleft", "fscore", "mcc")
shown <- c("metric", "rank", "sensitivity", "specificity", "precision")

knitr::kable(
  do.call(rbind, lapply(criteria, function(m) {
    best_cutoff(mdat, metric = m)[, shown]
  })),
  row.names = FALSE, digits = 3
)
```

Youden's J and the closest point to the top left corner are computed from
sensitivity and specificity alone. Both of those are conditioned on the true
class, so neither knows how rare the positives are, and both are free to buy
sensitivity with false positives that cost them nothing in the number being
maximized. `fscore` and `mcc` are computed from precision as well, so the
prevalence is in what they optimize.

| | |
|---|---|
| **Prevalence-blind** | `"youden"`, `"topleft"` |
| **Prevalence-aware** | `"fscore"`, `"mcc"`, `"cost"` |

The default is `"youden"`, because that is what `pROC::coords(x, "best")`
does and most people arrive expecting it. On imbalanced data it is the wrong
default, and [Balanced and imbalanced
data](howto-imbalanced-data.html#so-does-the-threshold) shows the two groups
landing nowhere near each other on the same model.

Any metric `evalmod()` calculates can be named, not only these five. The
direction is a property of the metric rather than an argument - `error`,
`cost`, `roc_dist`, the two error rates and the negative likelihood ratio are
minimized and the rest are maximized - so `metric = "accuracy"` means the
cutoff that maximizes accuracy without your having to say so.

Two of them are traps rather than criteria. `sensitivity` is largest when
everything is called positive and `specificity` when nothing is; both are the
correct answer to the question asked, and no warning says that it was
probably not the question meant.

## Saying what the mistakes cost

`"mcc"` is a reasonable default when you have nothing better to say. When you
do have something better to say, say it: `"cost"` minimizes
`cost_fp * FP + cost_fn * FN`, and the ratio between the two is the only part
that matters.

```{r}
costly <- lapply(c(1, 5, 20), function(fn) {
  cbind(
    cost_fn = fn,
    best_cutoff(
      mdat,
      metric = "cost", cost_fp = 1, cost_fn = fn
    )[, c("rank", "score", "sensitivity", "precision")]
  )
})

knitr::kable(do.call(rbind, costly), row.names = FALSE, digits = 3)
```

A missed positive costing twenty false alarms moves the threshold a long way
down, and it should. This is the criterion to reach for when the two mistakes
are genuinely different, because it is the only one that lets you write down
how different.

The first row is the warning that comes with it. Equal costs make the total
cost the error count, so minimizing it is maximizing accuracy - and where
positives are five percent of the data, calling almost nothing positive is
very nearly the most accurate thing a classifier can do. `cost_fp = cost_fn`
is not a neutral choice on imbalanced data. It is a strong claim, and usually
not the one intended.

## When scores are tied

A run of tied scores gets one row per instance in `metric_table()`, and by
default those rows split the run's positives and negatives between them. That
is what the curves need, and it is not what a threshold does: no threshold can
separate two instances that carry the same score.

`best_cutoff()` takes care of this. Rows sharing a `score` are one threshold
seen several times rather than several cutoffs, so the `rank` reported is the
last of the run - the number that `score >= ` that value actually calls
positive.

```{r}
binned <- round(scores)
tied <- best_cutoff(mmdata(binned, labels), metric = "mcc")

c(rank = tied$rank, applied = sum(binned >= tied$score))
```

`basic_ties = "hold"` gives every cutoff in a run the counts of the whole run,
which is the policy `pROC` and `ROCR` use. Pass it when you want [the table
underneath](howto-from-proc-rocr.html#tied-scores-produce-extra-rows) to line
up with theirs row for row. It rarely changes which cutoff comes back, because
the two policies agree at the edges of a run and that is where the optimum
tends to sit.

## Several models, several test sets

There is one row per model per test dataset, and nothing is averaged. A
cutoff belongs to the data it was read off.

```{r}
folds <- create_sim_samples(5, 20, 380, "good_er")
fdat <- mmdata(folds[["scores"]], folds[["labels"]],
  modnames = folds[["modnames"]], dsids = folds[["dsids"]]
)

knitr::kable(
  best_cutoff(fdat, metric = "mcc")[, c(
    "dsid", "value", "rank", "score", "precision"
  )],
  row.names = FALSE, digits = 3
)
```

Five test sets from the same generator, and the cutoffs do not agree. That
spread is the sampling noise in the choice, and it is worth looking at before
quoting any one of the numbers: on a small or imbalanced test set it is
usually larger than people expect. Averaging the `score` column is not
obviously the right summary either - across models the scores are not even on
a common scale - so the `rank`, or the value of the criterion, tends to travel
better.

## The number you quote will be optimistic

A threshold chosen on the same data the model is scored on is flattered by
however much the criterion was free to chase, in exactly the way a model
selected on its test set is. The fix is the ordinary one - choose the cutoff
on data the model has not seen, and report the metrics at that fixed cutoff
on data neither the model nor the cutoff has seen.

`precrec` does not do this for you. Nothing here resamples or cross-validates
the choice, and `best_cutoff()` will happily hand back the optimum of
whatever you give it. With [cross-validation
folds](howto-cross-validation.html) already in hand, choosing on some folds
and reporting on the rest is a few lines of ordinary R.

## Next

- [Balanced and imbalanced data](howto-imbalanced-data.html) - why the
  criterion matters more than it looks
- [Classification report](metrics-classification-report.html) - the metrics
  at a threshold you have already chosen
- [Get the numbers out](howto-results-as-data.html) - the table this takes a
  row of
