---
title: "Prepare your data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

For a single model on a single test set you do not need any of this - pass
`scores` and `labels` straight to `evalmod()`. This page is for everything
else.

```{r}
library(precrec)
```

## Four helpers

| Function | What it does |
|---|---|
| `join_scores()` | Collect the scores of several models into one list |
| `join_labels()` | Collect the labels of several test sets into one list |
| `mmdata()` | Turn those lists into the input `evalmod()` expects |
| `create_sim_samples()` | Make simulated data, for trying things out |

## Joining scores and labels

`join_scores()` accepts vectors, matrices and data frames, in any mixture,
and returns a list with one element per model.

```{r}
s1 <- c(1, 2, 3, 4)
s2 <- c(5, 6, 7, 8)

scores <- join_scores(s1, s2)
```

`join_labels()` does the same for observed labels.

```{r}
l1 <- c(1, 0, 1, 1)
l2 <- c(1, 0, 1, 0)

labels_same <- join_labels(l1, l1)
labels_diff <- join_labels(l1, l2)
```

Use the same label vector twice when two models were tested on the *same*
data, and two different ones when they were tested on different data. That
distinction is what tells `precrec` whether it is looking at several models
or several test sets.

## Building the input

`mmdata()` puts them together. Two identifiers decide how the result is read:
`modnames` names the models, `dsids` numbers the test sets.

```{r}
# Two models, one test set
mdat1 <- mmdata(scores, labels_same, modnames = c("mod1", "mod2"))

# One model, two test sets
mdat2 <- mmdata(scores, labels_diff, dsids = c(1, 2))
```

Leave them out and `precrec` uses sensible defaults. Set them when the
default guess is not what you meant.

## Simulated data

`create_sim_samples()` generates scores at a chosen quality level, which is
handy for experiments and for every example on this site.

| Level | Meaning |
|---|---|
| `random` | No better than chance |
| `poor_er` | Poor early retrieval |
| `good_er` | Good early retrieval |
| `excel` | Excellent |
| `perf` | Perfect |
| `all` | All five at once |

```{r}
# 10 test sets, 100 positives and 100 negatives, two quality levels
samps <- create_sim_samples(10, 100, 100, c("poor_er", "good_er"))

simdat <- mmdata(samps[["scores"]], samps[["labels"]],
  modnames = samps[["modnames"]], dsids = samps[["dsids"]]
)
```

## Missing and tied scores

`NA` scores are ranked last by default; `na_worst = FALSE` ranks them first.
Tied scores share a rank by default (`ties_method = "equiv"`); `"first"`
keeps the input order and `"random"` shuffles them.

## Next

- [Compare several models](howto-multiple-models.html)
- [Average over several test sets](howto-multiple-test-sets.html)
- [Evaluate cross-validation folds](howto-cross-validation.html)
