---
title: "Get the numbers out"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Every `precrec` object converts to a data frame, and every summary is a data
frame too. Nothing is locked inside the plotting code.

```{r}
library(precrec)

curves <- evalmod(scores = P10N10$scores, labels = P10N10$labels)
```

## What is in the object

`print()` summarizes the input and the results.

```{r}
curves
```

## The curve points

```{r}
df <- as.data.frame(curves)
head(df)
```

One row per supporting point, with the curve type and the model in their own
columns - the shape `ggplot2` and `dplyr` expect. `as.data.table()` returns
the same thing as a `data.table` when that package is installed.

## The summaries

| Function | Returns |
|---|---|
| `auc()` | Area under each curve, and its baseline |
| `pauc()` | Partial area, after `part()`, standardized either way |
| `auc_ci()` | Confidence interval of the area, over several test sets |
| `prbe()` | Precision-recall break-even point |
| `prob_metrics()` | Brier score, RMSE and log loss |

```{r}
knitr::kable(auc(curves))
```

Each returns a plain data frame, so subsetting is ordinary R.

```{r}
aucs <- auc(curves)
knitr::kable(subset(aucs, curvetypes == "PRC"))
```

## Every metric at every cutoff

`metric_table()` returns one row per cutoff and one column per metric. This
is the table `pROC::coords()` and `ROCR`'s cutoff slots give, and the shape
most work downstream of a curve needs.

```{r}
tab <- metric_table(scores = P10N10$scores, labels = P10N10$labels)

head(tab)
```

A row is the cutoff that calls the top `rank` instances positive, so `score`
is the score of the instance at that rank and the rule the row stands for is
`score >= that value`. `normalized_rank` is `rank / n`, the x axis of the
[basic metric plots](plots-basic-metrics.html). The first row calls nothing
positive, which is why its `score` and `label` are `NA`.

Its `precision` is `NA` too, and for a different reason: precision is
`TP / (TP + FP)`, and a rule that makes no positive predictions has no
denominator for it. The row that calls everything positive is missing its
`npv` at the other end, on the same argument. Every other metric on those
rows is measured - accuracy at rank 0 is the proportion of negatives, which
is exactly the point of
[Balanced and imbalanced data](howto-imbalanced-data.html).

Because it is a plain data frame, the question that usually follows is
ordinary R.

```{r}
best <- tab[which.max(tab$fscore), ]

knitr::kable(best[, c("rank", "score", "sensitivity", "precision", "fscore")])
```

`best_cutoff()` is that line with the parts that are easy to get wrong done
for you: which direction the metric is better in, which of several tied
cutoffs to return, and one row per model per test dataset rather than one
row overall.

```{r}
picked <- best_cutoff(
  scores = P10N10$scores, labels = P10N10$labels,
  metric = "fscore"
)

knitr::kable(
  picked[, c("metric", "value", "rank", "score", "sensitivity", "precision")]
)
```

It takes any of the metrics, plus `"youden"` and `"topleft"` for the two
criteria the cutpoint literature names rather than the metric table does.
Which one to optimize is a bigger decision than it looks, and the biggest
one when positives are rare - see
[Choose an operating point](howto-operating-point.html).

Metrics beyond the default fourteen come the same way they do from
`evalmod()`.

```{r}
lifted <- metric_table(
  scores = P10N10$scores, labels = P10N10$labels,
  metrics = c("lift", "jaccard")
)

head(lifted[, c("rank", "score", "precision", "lift", "jaccard")])
```

With several test datasets there is one block of rows per dataset and
nothing is averaged across them - a cutoff belongs to the dataset it was
read off. Build the object with `raw_curves = TRUE` if you pass one in
rather than letting `metric_table()` build it.

## The same metrics, long

`mode = "basic"` on its own gives the long form, which is what the plots
use.

```{r}
points <- evalmod(
  scores = P10N10$scores, labels = P10N10$labels,
  mode = "basic"
)

head(as.data.frame(points))
```

The `x` column is the normalized rank and `y` the value of the metric named
in `type`. It is the same data `metric_table()` returns, on its side. Reach
for this one to plot and for the table to decide.

An object that has already been calculated can be passed straight in, which
saves calculating it twice.

```{r}
identical(
  metric_table(points),
  metric_table(scores = P10N10$scores, labels = P10N10$labels)
)
```

## Feeding ggplot2 directly

`fortify()` is the `ggplot2` hook, so a `precrec` object can go straight into
`ggplot()`. See [customizing plots](plots-customizing.html).

## Next

- [Choose an operating point](howto-operating-point.html)
- [Metrics overview](metrics-overview.html)
- [Plots overview](plots-overview.html)
