---
title: "Coming from pROC or ROCR"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

A translation table. If you already have working `pROC` or `ROCR` code and
want the same numbers out of `precrec`, this page is the mapping.
[Comparison with other tools](metrics-other-tools.html) is the argument
about which to use; this one assumes you have decided.

Every equivalence below was checked against `pROC` 1.19.1 and `ROCR` 1.0.12.
Neither is a dependency of `precrec`, so their code is shown but not run when
this page is built - only the `precrec` chunks are.

```{r}
library(precrec)

set.seed(7)
scores <- c(rnorm(60, 1.1), rnorm(140, 0))
labels <- rep(c(1, 0), c(60, 140))
mdat <- mmdata(scores, labels)
```

## One object instead of one per question

This is the difference that shapes everything else. `pROC` builds a `roc`
object and asks it questions; `ROCR` builds a `prediction` and then a
`performance` object per metric pair. `precrec` calculates the curves and
the per-cutoff metrics once, and every function reads that result.

```r
# pROC: one object, one curve
r <- roc(labels, scores)
auc(r)
coords(r, "best")

# ROCR: a new performance object per question
p <- prediction(scores, labels)
performance(p, "tpr", "fpr")
performance(p, "prec", "rec")
performance(p, "auc")
```

```{r}
curves <- evalmod(mdat)

auc(curves)
```

One call, both curves, and the ROC and precision-recall areas together. The
practical consequence is that `mdat` and `curves` are worth keeping in a
variable: everything else on this page reads one of them.

## pROC

| `pROC` | `precrec` | |
|---|---|---|
| `roc(labels, scores)` | `evalmod(scores = , labels = )` | ROC and PRC together |
| `auc(r)` | `auc(curves)`, the `ROC` row | Identical |
| `ci.auc(r)` | `auc_ci(auc_delong(mdat))` | Identical - the same DeLong variance |
| `ci.auc(r, method = "bootstrap")` | `auc_ci(auc_boot(mdat))` | Percentile, and stratified |
| `roc.test(r1, r2)` | `auc_diff(auc_delong(mdat))` | Identical statistic and p-value |
| `roc.test(r1, r2, method = "bootstrap")` | `auc_diff(auc_boot(mdat))` | Paired on the same resamples |
| `coords(r, "best")` | `best_cutoff(mdat)` | Youden's J, the default in both |
| `coords(r, "best", best.method = "closest.topleft")` | `best_cutoff(mdat, metric = "topleft")` | |
| `coords(r, "all")` | `metric_table(mdat)` | Same rows, many more columns |
| `coords(r, t, input = "threshold")` | `classification_report(mdat, at = t)` | |
| `auc(r, partial.auc = c(1, 0.8), partial.auc.focus = "sp")` | `pauc(part(curves, xlim = c(0, 0.2)))` | Identical |
| `plot(r)` | `autoplot(curves, "ROC")` | See the axis note below |
| `multiclass.roc(labels, scores)` | `evalmod()` on more than two classes | [One-vs-rest](howto-multiclass.html) |
| `levels = `, `direction = ` | `posclass = ` | |

`roc.test()` is worth spelling out, because it is the reason many people
have `pROC` loaded at all:

```r
roc.test(roc(labels, a), roc(labels, b))   # DeLong, paired
```

```{r}
two <- mmdata(
  list(scores, scores * 0.6 + rnorm(200, 0, 0.9)),
  list(labels, labels),
  modnames = c("A", "B")
)

knitr::kable(auc_diff(auc_delong(two)), digits = 4)
```

`z_values` and `p_values` here are `roc.test()`'s `Z` and `p-value`. They
are the same statistic computed the same way, and on this example they
agree to every digit either package prints.

## ROCR

| `ROCR` | `precrec` | |
|---|---|---|
| `prediction(scores, labels)` | `mmdata(scores, labels)` | |
| `performance(p, "tpr", "fpr")` | `evalmod()`, the `ROC` curve | |
| `performance(p, "prec", "rec")` | `evalmod()`, the `PRC` curve | The numbers differ - see below |
| `performance(p, "auc")` | `auc(curves)`, the `ROC` row | Identical |
| `performance(p, "aucpr")` | `average_precision(curves)` | The step estimator; `auc()` interpolates instead |
| `performance(p, "acc")`, `"err"`, `"mat"`, ... | `metric_table(mdat)` | Identical, all at once |
| `performance(p, "prbe")` | `prbe(curves)` | |
| `performance(p, "cost", cost.fp = , cost.fn = )` | `metric_table(mdat, metrics = "cost", cost_fp = , cost_fn = )` | Identical |
| `performance(p, "fall")`, `"miss"`, `"rpp"` | The same names, as aliases | |
| `perf@x.values`, `perf@y.values` | `as.data.frame(curves)` | |
| `p@cutoffs` | The `score` column of `metric_table()` | |
| `prediction(list, list)`, `plot(perf, avg = )` | `mmdata(dsids = )`, then `evalmod()` | [Averaged by default](howto-multiple-test-sets.html) |

The one that replaces the most code is `metric_table()`, because `ROCR`
returns one metric per `performance()` call and this returns all of them:

```r
performance(p, "acc")@y.values[[1]]
performance(p, "mat")@y.values[[1]]   # and once more per metric
```

```{r}
tab <- metric_table(mdat)

head(tab[, c("rank", "score", "accuracy", "mcc", "precision", "sensitivity")])
```

The `score` column is `p@cutoffs` with one difference: `ROCR` puts `Inf`
first, for the cutoff that calls nothing positive, and `precrec` puts `NA`
there because no instance sits at rank `0`. The metrics on that row are the
same in both.

## Three things that will look wrong at first

### The precision-recall numbers do not match

They are not meant to. `ROCR` joins the raw precision-recall points with
straight lines; `precrec` interpolates between achievable points, which is
the reason the package exists. On the data above:

```{r}
areas <- auc(curves)

knitr::kable(
  data.frame(
    interpolated = areas$aucs[areas$curvetypes == "PRC"],
    step = average_precision(curves)$aps,
    ROCR_trapezoid = 0.6980509
  ),
  digits = 5
)
```

The gap widens as the positives get rarer, and
[Comparison with other tools](metrics-other-tools.html) shows it doing so.
If you need the step estimator - `scikit-learn`'s average precision, and
what most papers report under that name - it is `average_precision()`. What
`ROCR` draws has no counterpart here on purpose.

The ROC numbers, by contrast, agree exactly. If a ROC AUC has moved, it is
not the interpolation.

### Tied scores produce extra rows

`pROC` and `ROCR` have one row per distinct score. `precrec` has one row
per rank, so a run of tied scores gets a row per instance in the run, and
by default those rows split the run's positives and negatives evenly
between them. That is what the curves need, and it is not what a threshold
does.

```{r}
tied <- c(3, 3, 2, 2, 1, 1)
tied_labels <- c(1, 0, 1, 0, 1, 0)

held <- metric_table(
  scores = tied, labels = tied_labels, basic_ties = "hold"
)

knitr::kable(held[, c("rank", "score", "sensitivity", "specificity")])
```

`basic_ties = "hold"` gives every cutoff in a run the counts of the whole
run, which is the tie policy the other two packages use: the seven rows
above hold four distinct ones, and those four are the four rows of
`coords(r, "all")`. Pass it whenever the scores are rounded, binned or
voted - and pass it to `best_cutoff()` too, so the threshold it picks is
one a threshold can produce.

### The threshold is named differently

`coords(r, "best")` reports a midpoint between two neighboring scores.
`best_cutoff()` reports an observed score, and the rule is `score >= ` it.

```{r}
knitr::kable(
  best_cutoff(mdat)[, c("metric", "value", "rank", "score", "sensitivity")],
  digits = 4
)
```

The two name the same cutoff: every instance separates the same way, and
the sensitivity and specificity are identical. Only the number written down
as the threshold differs, and `precrec` uses a value the data actually
contains so that it can be applied without a rounding decision.

## The axes

`plot(roc)` puts specificity on the x axis and reverses it, so the curve
runs from the bottom right. `autoplot()` puts `1 - specificity` on it,
increasing left to right, which is the orientation the precision-recall
plot beside it uses.

Same curve, same area, different tick labels - and if the other one is what
your readers expect, `metric_curve()` will draw it:

```{r, fig.height = 4.5}
library(ggplot2)

sn_sp <- metric_curve(
  scores = scores, labels = labels,
  x_metric = "specificity", y_metric = "sensitivity"
)

autoplot(sn_sp) + scale_x_reverse()
```

## What has no equivalent

Honestly, and in the order they are missed:

- **Curve smoothing.** `smooth(roc)` has no counterpart, deliberately: a
  smoothed curve is a model of the data rather than a summary of it.
- **Venkatraman's test**, and tests at a fixed sensitivity or specificity.
  `auc_diff()` compares areas.
- **Unpaired comparison.** Both `auc_boot()` and `auc_delong()` require the
  models to be scored on the same instances.
- **Confidence bands from one test set.** `ci.se()` and `ci.sp()` have no
  counterpart; [confidence bands](plots-confidence-bands.html) here come
  from several test sets, and `auc_boot()` covers the single-sample case
  for the area rather than for the curve.
- **The standardized partial AUC** that `partial.auc.correct = TRUE`
  rescales to. `pauc()` reports the area itself.
- **Sample size calculation.** `power.roc.test()` has no counterpart.
- **`ROCR`'s convex hull, expected-cost curve and calibration error**
  (`"rch"`, `"ecost"`, `"cal"`), each a curve in a space of its own rather
  than a column of the metric table.
- **Coloring a curve by its cutoff.** `plot(perf, colorize = TRUE)` has no
  direct counterpart; [one metric against
  another](plots-metric-curve.html) is the nearest thing.

## Next

- [Comparison with other tools](metrics-other-tools.html) - which to use,
  and where the numbers differ
- [Get the numbers out](howto-results-as-data.html) - `metric_table()` and
  the long form
- [Balanced and imbalanced data](howto-imbalanced-data.html) - why the
  cutpoint criterion matters more than it looks
