---
title: "Use precrec with tidymodels"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

`precrec` needs no adapter. `tune::collect_predictions()` returns a fold
column, a truth column and one `.pred_<class>` column per class, which is the
shape `mmdata()` already takes through `nfold_df`.

```{r}
library(precrec)
```

## Set the positive class

This is the one thing to get right. `yardstick` treats the **first** factor
level as the event; `precrec` treats the **last** as the positive class. Pair
`yardstick`'s default score column with `precrec`'s default and the AUC comes
out inverted, with no warning.

```{r}
lv <- c("bad", "good")
truth <- factor(c("bad", "bad", "good", "good", "good"), levels = lv)
scores <- c(0.1, 0.2, 0.8, 0.9, 0.7)

c(
  default = auc(evalmod(mmdata(scores, truth)))$aucs[1],
  first = auc(evalmod(mmdata(scores, truth, posclass = "bad")))$aucs[1]
)
```

Feed `.pred_<last level>` and leave `posclass` alone, or feed
`.pred_<first level>` and name it in `posclass`.

## One model over folds

`collect_predictions()` gives `id` for the fold and `.row` for the instance.

```{r, eval = FALSE}
preds <- wf |>
  fit_resamples(folds, control = control_resamples(save_pred = TRUE)) |>
  collect_predictions()
```

There is no `tidymodels` on this machine, so the rest of the page runs
against a frame with the columns `collect_predictions()` returns.

```{r}
set.seed(1)
y <- sample(rep(c("good", "bad"), each = 40))
preds <- data.frame(
  id = rep(paste0("Fold", 1:4), each = 20), .row = 1:80,
  .pred_good = ifelse(y == "good", runif(80, 0.35, 1), runif(80, 0, 0.65)),
  truth = factor(y, levels = c("bad", "good"))
)
```

```{r}
curves <- evalmod(
  nfold_df = preds, score_cols = ".pred_good",
  lab_col = "truth", fold_col = "id",
  modnames = "rf", dsids = 1:4
)

knitr::kable(auc_ci(curves))
```

## Several models

`collect_predictions()` stacks models by `wflow_id` or `.config`, so widen
first and name one score column per model.

```{r, eval = FALSE}
wide <- preds |>
  tidyr::pivot_wider(
    id_cols = c(id, .row, truth),
    names_from = wflow_id, values_from = .pred_good
  )

evalmod(
  nfold_df = as.data.frame(wide), score_cols = c("rec_rf", "rec_glm"),
  lab_col = "truth", fold_col = "id",
  modnames = c("rec_rf", "rec_glm"), dsids = 1:4
)
```

## More than two classes

Hand the `.pred_` columns over as a matrix. Classes are matched by column
name, so drop the prefix.

```{r, eval = FALSE}
sc <- as.matrix(preds[grep("^\\.pred_", names(preds))])
colnames(sc) <- sub("^\\.pred_", "", colnames(sc))

evalmod(mmdata(sc, preds$truth))
```

One-vs-rest cannot be combined with `nfold_df`, so pool the folds for a
multiclass run, or loop over them yourself.

## The report

`.pred_class` is the hard prediction `parsnip` derives at a 0.5 threshold,
and `classification_report()` takes that threshold directly.

```{r}
classification_report(mmdata(preds$.pred_good, preds$truth), at = 0.5)
```

## Next

- [Evaluate cross-validation folds](howto-cross-validation.html)
- [How precrec compares with other tools](metrics-other-tools.html)
