---
title: "Balanced and imbalanced data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

This is the reason the package exists. When positives are rare, a ROC curve
can look excellent for a classifier that is not usable, while the
precision-recall curve shows the problem plainly.

```{r}
library(precrec)
library(ggplot2)
```

## Two datasets, same models

The only difference between them is the class balance: 100 positives against
100 negatives, and 25 positives against 100 negatives.

```{r}
balanced <- create_sim_samples(20, 100, 100, "all")
imbalanced <- create_sim_samples(20, 25, 100, "all")

bdat <- mmdata(balanced[["scores"]], balanced[["labels"]],
  modnames = balanced[["modnames"]], dsids = balanced[["dsids"]]
)
idat <- mmdata(imbalanced[["scores"]], imbalanced[["labels"]],
  modnames = imbalanced[["modnames"]], dsids = imbalanced[["dsids"]]
)
```

## Balanced

```{r}
autoplot(evalmod(bdat))
```

## Imbalanced

```{r}
autoplot(evalmod(idat))
```

## What changed

The ROC curves are almost identical between the two. The precision-recall
curves are not: every model drops, because precision depends on how many
negatives there are to be mistaken for positives, and the ROC axes do not.

A ROC curve is therefore the wrong plot for reporting how a classifier will
behave on data where positives are rare - which is most screening, detection
and diagnostic problems.

## The baseline moves too

A precision-recall curve is read against its baseline, which sits at the
proportion of positives: 0.5 above, 0.2 in the imbalanced case. The same
curve means something different against a different baseline, which is why
the two plots above cannot be compared by eye alone.

## Further reading

- [The Precision-Recall Plot Is More Informative than the ROC Plot When
  Evaluating Binary Classifiers on Imbalanced
  Datasets](https://doi.org/10.1371/journal.pone.0118432) - the paper behind
  this package
- [Classifier evaluation with imbalanced
  datasets](https://classeval.wordpress.com/) - a companion site with
  practical tips

## Next

- [Evaluate more than two classes](howto-multiclass.html)
- [AUC and other curve summaries](metrics-auc.html)
