---
title: "Prepare your data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

For a single model on a single test set you do not need any of this - pass
`scores` and `labels` straight to `evalmod()`. This page is for everything
else.

```{r}
library(precrec)
```

## Four helpers

| Function | What it does |
|---|---|
| `join_scores()` | Collect the scores of several models into one list |
| `join_labels()` | Collect the labels of several test sets into one list |
| `mmdata()` | Turn those lists into the input `evalmod()` expects |
| `create_sim_samples()` | Make simulated data, for trying things out |

## Joining scores and labels

`join_scores()` accepts vectors, matrices and data frames, in any mixture,
and returns a list with one element per model.

```{r}
s1 <- c(1, 2, 3, 4)
s2 <- c(5, 6, 7, 8)

scores <- join_scores(s1, s2)
```

`join_labels()` does the same for observed labels.

```{r}
l1 <- c(1, 0, 1, 1)
l2 <- c(1, 0, 1, 0)

labels_same <- join_labels(l1, l1)
labels_diff <- join_labels(l1, l2)
```

Use the same label vector twice when two models were tested on the *same*
data, and two different ones when they were tested on different data. That
distinction is what tells `precrec` whether it is looking at several models
or several test sets.

## Building the input

`mmdata()` puts them together. Two identifiers decide how the result is read:
`modnames` names the models, `dsids` numbers the test sets.

```{r}
# Two models, one test set
mdat1 <- mmdata(scores, labels_same, modnames = c("mod1", "mod2"))

# One model, two test sets
mdat2 <- mmdata(scores, labels_diff, dsids = c(1, 2))
```

Leave them out and `precrec` uses sensible defaults. Set them when the
default guess is not what you meant.

## Simulated data

`create_sim_samples()` generates scores at a chosen quality level, which is
handy for experiments and for every example on this site.

| Level | Meaning |
|---|---|
| `random` | No better than chance |
| `poor_er` | Poor early retrieval |
| `good_er` | Good early retrieval |
| `excel` | Excellent |
| `perf` | Perfect |
| `all` | All five at once |

```{r}
# 10 test sets, 100 positives and 100 negatives, two quality levels
samps <- create_sim_samples(10, 100, 100, c("poor_er", "good_er"))

simdat <- mmdata(samps[["scores"]], samps[["labels"]],
  modnames = samps[["modnames"]], dsids = samps[["dsids"]]
)
```

## Missing and tied scores

`NA` scores are ranked last by default; `na_worst = FALSE` ranks them first.
Tied scores share a rank by default (`ties_method = "equiv"`); `"first"`
keeps the input order and `"random"` shuffles them.

## Which direction the score runs

`precrec` ranks the highest score first, so a score has to be larger for
instances that are more likely to be positive. Plenty of quantities run the
other way - a p-value from Fisher's exact test, a distance to a cluster
center, an error term - and for those, negate the score:

```{r}
pvals <- c(0.001, 0.02, 0.3, 0.4, 0.8, 0.9)
obs <- c(1, 1, 0, 1, 0, 0)

evalmod(scores = -pvals, labels = obs)
```

There is no argument for this. Negation is the whole of it, and it changes
nothing else: the curves, the areas and the basic metrics all read the
ranking, and reversing the score reverses the ranking exactly.

The one thing negation does not survive is `prob_metrics()`. The Brier score
and the log loss read the *values* of the scores rather than their order, so
they need genuine probabilities on 0 to 1 - `-pvals` is rejected. Convert
rather than negate when you want those: `1 - pvals` ranks the same way and
stays in range, though whether it is calibrated is a separate question.

## What cannot be used as input

`precrec` always needs the per-instance scores and labels. Precision and
recall already calculated at a handful of thresholds cannot be used, and
neither can a table of true and false positive rates.

This is a real limit rather than a missing feature. `precrec` draws the
precision-recall curve by interpolating *between consecutive ranks*, and
that interpolation needs to know how many positives and negatives lie
between one point and the next - which is exactly what a table of summary
points has already discarded. The same ranking is what the areas, the
confidence bands and the basic metrics are calculated from.

If the summary points are genuinely all you have, they are still a line, and
`ggplot2` will draw it. What they are not is a `precrec` object, and the
accuracy `precrec` exists for is not available from them.

## Next

- [Compare several models](howto-multiple-models.html)
- [Average over several test sets](howto-multiple-test-sets.html)
- [Evaluate cross-validation folds](howto-cross-validation.html)
