---
title: "Balanced and imbalanced data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

This is the reason the package exists. When positives are rare, a ROC curve
can look excellent for a classifier that is not usable, while the
precision-recall curve shows the problem plainly.

```{r}
library(precrec)
library(ggplot2)
```

## Two datasets, same models

The only difference between them is the class balance: 100 positives against
100 negatives, and 25 positives against 100 negatives.

```{r}
balanced <- create_sim_samples(20, 100, 100, "all")
imbalanced <- create_sim_samples(20, 25, 100, "all")

bdat <- mmdata(balanced[["scores"]], balanced[["labels"]],
  modnames = balanced[["modnames"]], dsids = balanced[["dsids"]]
)
idat <- mmdata(imbalanced[["scores"]], imbalanced[["labels"]],
  modnames = imbalanced[["modnames"]], dsids = imbalanced[["dsids"]]
)
```

## Balanced

```{r}
autoplot(evalmod(bdat))
```

## Imbalanced

```{r}
autoplot(evalmod(idat))
```

## What changed

The ROC curves are almost identical between the two. The precision-recall
curves are not: every model drops, because precision depends on how many
negatives there are to be mistaken for positives, and the ROC axes do not.

A ROC curve is therefore the wrong plot for reporting how a classifier will
behave on data where positives are rare - which is most screening, detection
and diagnostic problems.

## The baseline moves too

A precision-recall curve is read against its baseline, which sits at the
proportion of positives: 0.5 above, 0.2 in the imbalanced case. The same
curve means something different against a different baseline, which is why
the two plots above cannot be compared by eye alone.

## So does the threshold

Everything above is about reading a curve. The last step of most real work
is picking a point on it, and the criterion that picks it is subject to the
same problem.

Twenty-five positives in five hundred, and four of the criteria
[best_cutoff()](howto-operating-point.html) offers, on the same model:

```{r}
set.seed(42)
rare <- create_sim_samples(1, 25, 475, "good_er")
rdat <- mmdata(rare[["scores"]], rare[["labels"]])

shown <- c(
  "metric", "rank", "score", "sensitivity", "precision", "fscore", "mcc"
)
picks <- lapply(
  c("youden", "topleft", "fscore", "mcc"),
  function(criterion) best_cutoff(rdat, metric = criterion)[, shown]
)

knitr::kable(do.call(rbind, picks), row.names = FALSE, digits = 3)
```

Four criteria, three different thresholds, and the disagreement is not a
rounding detail: Youden's J calls 78 of the 500 instances positive and
`mcc` calls 22. The model is the same one and the data is the same data.

Youden's J and the closest point to the top left corner are both computed
from sensitivity and specificity, and both of those are conditioned on the
true class, so neither knows that positives are rare. They are free to buy
sensitivity with false positives that cost them nothing, and they do -
precision at the cutoff Youden's J picks is half of what it is at the
one `mcc` picks. `fscore` and `mcc` are computed from precision as well, so the
prevalence is in the number they are maximizing.

That does not make `mcc` the right answer either. It buys its precision
with sensitivity, and whether that trade is the one you want is a question
about the application, not about the data. What the table settles is
narrower and more useful: **the criterion is a choice**, and on imbalanced
data the prevalence-blind ones and the prevalence-aware ones do not land
anywhere near each other.

One caveat that applies whatever the balance: a threshold chosen on the same
data the model was evaluated on is optimistic, by however much the criterion
was free to chase. If the number is going to be quoted, choose it on data
the model has not seen.

## Further reading

- [The Precision-Recall Plot Is More Informative than the ROC Plot When
  Evaluating Binary Classifiers on Imbalanced
  Datasets](https://doi.org/10.1371/journal.pone.0118432) - the paper behind
  this package
- [Classifier evaluation with imbalanced
  datasets](https://classeval.wordpress.com/) - a companion site with
  practical tips

## Next

- [Choose an operating point](howto-operating-point.html)
- [Evaluate more than two classes](howto-multiclass.html)
- [AUC and other curve summaries](metrics-auc.html)
