---
title: "Balanced and imbalanced data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

This is the reason the package exists. When positives are rare, a ROC curve
can look excellent for a classifier that is not usable, while the
precision-recall curve shows the problem plainly.

```{r}
library(precrec)
library(ggplot2)
```

## Two datasets, same models

The only difference between them is the class balance: 100 positives against
100 negatives, and 25 positives against 100 negatives.

```{r}
balanced <- create_sim_samples(20, 100, 100, "all")
imbalanced <- create_sim_samples(20, 25, 100, "all")

bdat <- mmdata(balanced[["scores"]], balanced[["labels"]],
  modnames = balanced[["modnames"]], dsids = balanced[["dsids"]]
)
idat <- mmdata(imbalanced[["scores"]], imbalanced[["labels"]],
  modnames = imbalanced[["modnames"]], dsids = imbalanced[["dsids"]]
)
```

## Balanced

```{r}
autoplot(evalmod(bdat))
```

## Imbalanced

```{r}
autoplot(evalmod(idat))
```

## What changed

The ROC curves are almost identical between the two. The precision-recall
curves are not: every model drops, because precision depends on how many
negatives there are to be mistaken for positives, and the ROC axes do not.

A ROC curve is therefore the wrong plot for reporting how a classifier will
behave on data where positives are rare - which is most screening, detection
and diagnostic problems.

## The baseline moves too

A precision-recall curve is read against its baseline, which sits at the
proportion of positives: 0.5 above, 0.2 in the imbalanced case. The same
curve means something different against a different baseline, which is why
the two plots above cannot be compared by eye alone.

The same is true of the area, and there the baseline is easier to forget
because the plot is not in front of you. `auc()` reports it in the
`baselines` column:

```{r}
areas <- rbind(auc(evalmod(bdat)), auc(evalmod(idat)))

knitr::kable(
  subset(areas, curvetypes == "PRC" & modnames == "good_er" & dsids == 1),
  row.names = FALSE, digits = 3
)
```

The same model on the two datasets above. The area falls, and so does the
baseline it is read against - from 0.5 to 0.2. How much of the fall is the
classifier doing worse and how much is the class balance moving underneath
it is not a question the area answers on its own, which is why the two
numbers belong together.

## Why not fold the baseline in?

If the area has to be read against its baseline, the obvious thought is to
build the baseline into the area and be done with it. On the ROC side there
is such a thing - the McClish correction, `pauc(corrected = TRUE)`, which
rescales a partial area between chance and perfect.

For the precision-recall curve the published proposal is
Precision-Recall-Gain (Flach & Kull 2015). It rescales both axes against the
baseline `pi`,

```
precG = 1 - (pi / (1 - pi)) * ((1 - prec) / prec)
recG  = 1 - (pi / (1 - pi)) * ((1 - rec)  / rec)
```

so that chance sits at `0` and perfect at `1` whatever the class balance,
and the area under the resulting curve (AUPRG) is one number with the
baseline already in it. The argument behind it is a serious one, and it
depends on the curve being interpolated correctly, which is this package's
whole business.

It does not do what someone reaching for it wants. Here the same 200
positives appear in every row and negatives are added underneath them, so
the classifier and the positive sample never change - only the prevalence
does.

```{r}
gain <- function(v, pi) 1 - (pi / (1 - pi)) * ((1 - v) / v)

auprg <- function(scores, labels) {
  pi <- mean(labels == 1)
  d <- as.data.frame(evalmod(scores = scores, labels = labels))
  prc <- d[d$type == "PRC", ]
  rg <- gain(prc$x, pi)
  pg <- gain(prc$y, pi)
  ok <- is.finite(rg) & is.finite(pg)
  # Integrate precG over recG from 0 to 1
  grid <- sort(unique(c(0, 1, rg[ok][rg[ok] > 0 & rg[ok] < 1])))
  v <- approx(rg[ok], pg[ok], xout = grid, rule = 2, ties = "ordered")$y
  sum(diff(grid) * (head(v, -1) + tail(v, -1)) / 2)
}

set.seed(11)
pos <- rnorm(200, 1.2)
negpool <- rnorm(9800, 0)

row_at <- function(nn) {
  scores <- c(pos, negpool[seq_len(nn)])
  labels <- rep(c(1, 0), c(200, nn))
  areas <- auc(evalmod(scores = scores, labels = labels))

  data.frame(
    positives = sprintf("%.0f%%", 100 * 200 / (200 + nn)),
    roc = areas$aucs[areas$curvetypes == "ROC"],
    prc = areas$aucs[areas$curvetypes == "PRC"],
    prc_baseline = areas$baselines[areas$curvetypes == "PRC"],
    auprg = auprg(scores, labels)
  )
}

knitr::kable(
  do.call(rbind, lapply(c(200, 1800, 9800), row_at)),
  row.names = FALSE, digits = 3
)
```

On the AUPRG scale a coin flip scores `0` and a perfect ranking `1`, so
`0.639` and `0.941` are far apart on it. The ROC area holds within 0.01, as
it should. The precision-recall area falls, which is the whole point of the
page. And AUPRG **rises** - three tenths of its range, in the opposite
direction, on a classifier that did not change.

That is the thing to notice. Folding the baseline in did not remove the
dependence on prevalence; it reversed it. A number that says a classifier
got better because its positives got rarer is not more comparable across
datasets than the one it replaced - it is incomparable in a direction that
happens to flatter the hard case. The `baselines` column is less clever and
harder to misread.

So `precrec` reports the area and the baseline as two numbers and leaves
them that way, and there is no `auprg()` here. The few lines above are all
it takes if you want it for a curve you have already computed, and Flach &
Kull's own argument for it - that the precision-recall area is a poorly
founded summary quite apart from any question of class balance - is worth
reading on its own terms.

## The baseline is an asymptote

The prevalence is what the area is worth by chance *in the limit*. An area
measured on a finite sample scatters around its chance level rather than
sitting on it, and how widely depends on the number of positives - which is
what imbalance takes away.

Here is a classifier with no signal whatever, scored four hundred times on
fresh random labels. Twenty positives in a thousand, the same 2% as above.
The column is the area divided by the baseline `auc()` reports for it, so
chance is 1.

```{r}
np <- 20
nn <- 980
prevalence <- np / (np + nn)

noise <- t(sapply(seq_len(400), function(i) {
  set.seed(i)
  curves <- evalmod(
    scores = rnorm(np + nn), labels = rep(c(1, 0), c(np, nn))
  )
  areas <- auc(curves)
  top <- pauc(part(curves, xlim = c(0, 0.1)))

  c(
    whole = areas$aucs[areas$curvetypes == "PRC"],
    top = top$spaucs[top$curvetypes == "PRC"]
  ) / prevalence
}))

knitr::kable(
  data.frame(
    region = c("the whole curve", "the first tenth of recall"),
    mean = colMeans(noise),
    median = apply(noise, 2, median),
    over_2x = colMeans(noise > 2)
  ),
  row.names = FALSE, digits = 2
)
```

A classifier that knows nothing averages 1.15 times its baseline over the
whole curve, and 1.86 times over the top of the ranking - where it clears
twice the baseline about one run in five. The median tells you what kind of
error this is: at 0.89 it sits *below* chance, so the distribution is skewed
rather than shifted. Most draws land near or under the baseline and a few
land far above it, and any one dataset is one draw.

Two hundred positives brings the first figure to 1.01 and two thousand
removes it entirely, so what drives this is the number of positives rather
than the balance. `average_precision()` shows the same thing, so it is not
an artifact of the interpolation either. Rare positives cost you twice over:
few positives to estimate from, and a small baseline with room above it for
the tail to run into.

None of this makes the baseline the wrong number - it is the right one, and
the fall from 0.5 to 0.2 above is real. What it means is that dividing an
area by its baseline is not a test.

## So put an interval on it

`auc_boot()` resamples a single test set, and `auc_ci()` reads an interval
off the resamples. Both now report the baseline beside the area, so the
comparison is one table rather than two.

```{r}
set.seed(57)
pure_noise <- auc_boot(
  scores = rnorm(np + nn), labels = rep(c(1, 0), c(np, nn)),
  boot_n = 500, seed = 1
)

knitr::kable(
  auc_ci(pure_noise)[, c(
    "curvetypes", "aucs", "baselines", "lower_bound", "upper_bound"
  )],
  row.names = FALSE, digits = 3
)
```

The ROC area lands on 0.5, as it should for a classifier built out of
`rnorm()`. The precision-recall area is 1.8 times its baseline, which read on
its own is the sort of number people write down. The interval contains the
baseline, which is the correct answer.

The interval is well behaved on this case even at twenty positives: over a
hundred and twenty runs of the simulation above, the 95% bootstrap interval
covered the baseline 94% of the time and its lower bound cleared the
baseline 2.5% of the time, which is what a 95% interval is supposed to do.
The resampling is stratified, so every resample holds the same class balance
as the test set and the baseline does not move underneath the quantity being
estimated.

With several test sets, `auc_ci()` does the same job from the variation
between them, and averages the baseline over the same datasets it averaged
the area over - a fold need not hold the classes in the proportions the
whole dataset does.

`pauc()` carries the same two numbers for a restricted region, where they
are easier still to get wrong: a standardized partial ROC area over false
positive rates up to 0.2 has a chance level of 0.1, not 0.5.

## So does the threshold

Everything above is about reading a curve. The last step of most real work
is picking a point on it, and the criterion that picks it is subject to the
same problem.

Twenty-five positives in five hundred, and four of the criteria
[best_cutoff()](howto-operating-point.html) offers, on the same model:

```{r}
set.seed(42)
rare <- create_sim_samples(1, 25, 475, "good_er")
rdat <- mmdata(rare[["scores"]], rare[["labels"]])

shown <- c(
  "metric", "rank", "score", "sensitivity", "precision", "fscore", "mcc"
)
picks <- lapply(
  c("youden", "topleft", "fscore", "mcc"),
  function(criterion) best_cutoff(rdat, metric = criterion)[, shown]
)

knitr::kable(do.call(rbind, picks), row.names = FALSE, digits = 3)
```

Four criteria, three different thresholds, and the disagreement is not a
rounding detail: Youden's J calls 78 of the 500 instances positive and
`mcc` calls 22. The model is the same one and the data is the same data.

Youden's J and the closest point to the top left corner are both computed
from sensitivity and specificity, and both of those are conditioned on the
true class, so neither knows that positives are rare. They are free to buy
sensitivity with false positives that cost them nothing, and they do -
precision at the cutoff Youden's J picks is half of what it is at the
one `mcc` picks. `fscore` and `mcc` are computed from precision as well, so the
prevalence is in the number they are maximizing.

That does not make `mcc` the right answer either. It buys its precision
with sensitivity, and whether that trade is the one you want is a question
about the application, not about the data. What the table settles is
narrower and more useful: **the criterion is a choice**, and on imbalanced
data the prevalence-blind ones and the prevalence-aware ones do not land
anywhere near each other.

One caveat that applies whatever the balance: a threshold chosen on the same
data the model was evaluated on is optimistic, by however much the criterion
was free to chase. If the number is going to be quoted, choose it on data
the model has not seen.

## Further reading

- [The Precision-Recall Plot Is More Informative than the ROC Plot When
  Evaluating Binary Classifiers on Imbalanced
  Datasets](https://doi.org/10.1371/journal.pone.0118432) - the paper behind
  this package
- [Precision-Recall-Gain Curves: PR Analysis Done
  Right](https://papers.nips.cc/paper/5867-precision-recall-gain-curves-pr-analysis-done-right)
  - Flach & Kull's case against the precision-recall area, and the
  alternative measured above
- [Classifier evaluation with imbalanced
  datasets](https://classeval.wordpress.com/) - a companion site with
  practical tips

## Next

- [Choose an operating point](howto-operating-point.html)
- [Evaluate more than two classes](howto-multiclass.html)
- [AUC and other curve summaries](metrics-auc.html)
