---
title: "AUC and other curve summaries"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

These summarize a whole curve in one number, so they do not depend on picking
a cutoff.

```{r}
library(precrec)
library(ggplot2)

curves <- evalmod(scores = P10N10$scores, labels = P10N10$labels)
```

## Area under the curve

```{r}
knitr::kable(auc(curves))
```

| Curve | What the area means | Baseline |
|---|---|---|
| ROC | Probability a random positive outranks a random negative | 0.5 |
| PRC | Average precision over all recall levels | The proportion of positives |

The ROC baseline is always 0.5. The precision-recall baseline is not - it
moves with the class balance, so a PRC area of 0.4 is poor on balanced data
and good when positives are 5% of the total.

`precrec` computes the precision-recall area from the properly interpolated
curve. Tools that join raw points with straight lines overestimate it.

## Partial areas

When only part of the curve matters - the low-false-positive end, say -
`part()` restricts the range and `pauc()` reports the area over it.

```{r}
partial <- part(curves, xlim = c(0, 0.25))

knitr::kable(pauc(partial))
```

`spaucs` in that table are standardized partial AUCs, rescaled to 0 to 1 so
that ranges of different widths can be compared. Plotting the restricted
object draws the partial curve; see [partial
curves](plots-partial-curves.html).

## Confidence intervals

With several test sets, `auc_ci()` puts an interval around the mean area.

```{r}
samps <- create_sim_samples(10, 100, 100, "good_er")
mdat <- mmdata(samps[["scores"]], samps[["labels"]], dsids = samps[["dsids"]])
mcurves <- evalmod(mdat)

knitr::kable(auc_ci(mcurves))
```

`alpha` sets the level - `alpha = 0.01` for 99% - and `dtype = "t"` uses the
t distribution instead of the normal, which is the better choice for a small
number of folds.

```{r}
knitr::kable(auc_ci(mcurves, alpha = 0.01, dtype = "t"))
```

## Break-even point

The break-even point is where precision and recall are equal - the one point
on the precision-recall curve where the two are in balance.

`prbe()` reads the individual curves, so ask `evalmod()` to keep them.

```{r}
raw <- evalmod(mdat, raw_curves = TRUE)

knitr::kable(prbe(raw))
```

A curve can cross the diagonal more than once, and then the result holds one
row per crossing. A curve that never reaches equal precision and recall gets
a single row of `NA`.

`prbe()` reads the crossing off the interpolated curve. Finding it by linear
interpolation between raw precision-recall points, as some tools do, is the
error this package exists to avoid.

## The fast path

If the ROC area is all you need, `mode = "aucroc"` computes it from the U
statistic of the Mann-Whitney test without building the curve at all. See
[working with large datasets](howto-large-datasets.html).

```{r}
knitr::kable(as.data.frame(
  evalmod(scores = P10N10$scores, labels = P10N10$labels, mode = "aucroc")
))
```

## Next

- [Probability-based metrics](measures-probability.html)
- [Partial curves](plots-partial-curves.html)
