---
title: "AUC and other curve summaries"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

These summarize a whole curve in one number, so they do not depend on picking
a cutoff.

```{r}
library(precrec)
library(ggplot2)

curves <- evalmod(scores = P10N10$scores, labels = P10N10$labels)
```

## Area under the curve

```{r}
knitr::kable(auc(curves))
```

| Curve | What the area means | Baseline |
|---|---|---|
| ROC | Probability a random positive outranks a random negative | 0.5 |
| PRC | Mean precision across all recall levels | The proportion of positives |

The ROC baseline is always 0.5. The precision-recall baseline is not - it
moves with the class balance, so a PRC area of 0.4 is poor on balanced data
and good when positives are 5% of the total.

`precrec` computes the precision-recall area from the properly interpolated
curve. Tools that join raw points with straight lines overestimate it.

## Average precision

Average precision is the other way to summarize a precision-recall curve: the
precision at each cutoff, weighted by the recall it gains over the cutoff
before it.

```{r}
knitr::kable(average_precision(curves))
```

It is a different estimator from the area above, not a different way of
adding up the same one. It joins the raw points with horizontal steps instead
of interpolating between them, so it reads high wherever the two disagree -
which on this dataset is by about 0.006.

```{r}
knitr::kable(auc(curves))
```

Prefer the area under the interpolated curve. Average precision is here
because several other packages report it under this name, and because the
size of the gap between the two is worth being able to see.

## Partial areas

When only part of the curve matters - the low-false-positive end, say -
`part()` restricts the range and `pauc()` reports the area over it.

```{r}
partial <- part(curves, xlim = c(0, 0.25))

knitr::kable(pauc(partial))
```

`spaucs` in that table are standardized partial AUCs, rescaled to 0 to 1 so
that ranges of different widths can be compared. Plotting the restricted
object draws the partial curve; see [partial
curves](plots-partial-curves.html).

## Confidence intervals

With several test sets, `auc_ci()` puts an interval around the mean area.

```{r}
samps <- create_sim_samples(10, 100, 100, "good_er")
mdat <- mmdata(samps[["scores"]], samps[["labels"]], dsids = samps[["dsids"]])
mcurves <- evalmod(mdat)

knitr::kable(auc_ci(mcurves))
```

`alpha` sets the level - `alpha = 0.01` for 99% - and `dtype = "t"` uses the
t distribution instead of the normal, which is the better choice for a small
number of folds.

```{r}
knitr::kable(auc_ci(mcurves, alpha = 0.01, dtype = "t"))
```

## Break-even point

The break-even point is where precision and recall are equal - the one point
on the precision-recall curve where the two are in balance.

`prbe()` reads the individual curves, so ask `evalmod()` to keep them.

```{r}
raw <- evalmod(mdat, raw_curves = TRUE)

knitr::kable(prbe(raw))
```

A curve can cross the diagonal more than once, and then the result holds one
row per crossing. A curve that never reaches equal precision and recall gets
a single row of `NA`.

`prbe()` reads the crossing off the interpolated curve. Finding it by linear
interpolation between raw precision-recall points, as some tools do, is the
error this package exists to avoid.

## Averaging over classes

For a multiclass evaluation, `auc()` adds the macro-average of the per-class
areas. `macro_weight` chooses how the classes are weighted.

```{r}
mccurves <- evalmod(scores = C3N150$scores, labels = C3N150$labels)

knitr::kable(auc(mccurves))
```

The default, `"uniform"`, gives every class the same weight, so a rare class
counts as much as a common one. `"prevalence"` weights each class by the
number of observations it has, and the rows are named
`macro-average-weighted` to keep the two apart.

```{r}
knitr::kable(auc(mccurves, macro_weight = "prevalence"))
```

The two agree on a balanced dataset, as this one is. Which to use is the same
question as always: uniform if every class matters equally, prevalence if you
care about the average case. Other packages call these two `roc_aunu` and
`roc_aunp`. See [more than two classes](howto-multiclass.html).

## The fast path

If the ROC area is all you need, `mode = "aucroc"` computes it from the U
statistic of the Mann-Whitney test without building the curve at all. See
[working with large datasets](howto-large-datasets.html).

```{r}
knitr::kable(as.data.frame(
  evalmod(scores = P10N10$scores, labels = P10N10$labels, mode = "aucroc")
))
```

## Next

- [Probability-based metrics](measures-probability.html)
- [Partial curves](plots-partial-curves.html)
