---
title: "Confusion-matrix rates"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Everything here is a ratio of two cells of the 2x2 table, evaluated at every
cutoff.

|  | Predicted positive | Predicted negative |
|---|---|---|
| **Actually positive** | TP | FN |
| **Actually negative** | FP | TN |

```{r}
library(precrec)
library(ggplot2)

points <- evalmod(scores = P10N10$scores, labels = P10N10$labels,
  mode = "basic"
)
```

## Overall

| Measure | Formula | Range |
|---|---|---|
| `accuracy` | (TP + TN) / all | 0 to 1 |
| `error` | (FP + FN) / all | 0 to 1 |

Accuracy is the measure to distrust first. On data with 1% positives, calling
everything negative scores 0.99.

```{r}
autoplot(points, c("accuracy", "error"))
```

## Rates over the actual classes

These divide by a row of the table, so they do not move when the class
balance changes.

| Measure | Formula | Also known as |
|---|---|---|
| `sensitivity` | TP / (TP + FN) | recall, TPR, hit rate |
| `specificity` | TN / (TN + FP) | TNR, selectivity |
| `fpr` | FP / (FP + TN) | fall-out, 1 - specificity |
| `fnr` | FN / (TP + FN) | miss rate, 1 - sensitivity |

```{r}
autoplot(points, c("sensitivity", "specificity"))
```

The ROC curve is `sensitivity` against `fpr`, which is why it is blind to
class balance - both axes are row-wise rates.

## Rates over the predicted classes

These divide by a column, so they *do* move with the class balance. That is
what makes them informative on imbalanced data, and what makes them
impossible to transfer between datasets.

| Measure | Formula | Also known as |
|---|---|---|
| `precision` | TP / (TP + FP) | PPV |
| `npv` | TN / (TN + FN) | negative predictive value |
| `false_discovery_rate` | FP / (TP + FP) | 1 - precision |
| `false_omission_rate` | FN / (TN + FN) | 1 - NPV |

```{r}
autoplot(points, c("precision", "npv"))
```

The precision-recall curve is `precision` against `sensitivity` - one
column-wise rate against one row-wise rate. See [balanced and imbalanced
data](howto-imbalanced-data.html).

## How much gets flagged

| Measure | Formula | Also known as |
|---|---|---|
| `predicted_positive_rate` | (TP + FP) / all | rate of positive predictions |
| `predicted_negative_rate` | (TN + FN) / all | rate of negative predictions |

Useful when the cost of review is what constrains you: they say how much work
a cutoff creates, regardless of whether the work is well spent.

```{r}
ppr <- evalmod(scores = P10N10$scores, labels = P10N10$labels,
  mode = "basic", metrics = "predicted_positive_rate"
)

autoplot(ppr, "predicted_positive_rate")
```

## Undefined ends

Precision has no value at the cutoff where nothing is predicted positive, and
NPV none where nothing is predicted negative. `precrec` fills each in from
its neighbor rather than dropping the point.

## Next

- [Agreement and balance](measures-agreement.html)
- [Ranking and cost](measures-ranking-cost.html)
