---
title: "Probability-based metrics"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Everything else in `precrec` reads the *order* of the scores. These three
read their *values*, and ask a different question: not "does the classifier
rank well" but "are its probabilities honest".

```{r}
library(precrec)

# Scores drawn from beta distributions, so they are probabilities
samps <- create_sim_samples(4, 100, 100, c("poor_er", "good_er"))

mdat <- mmdata(samps[["scores"]], samps[["labels"]],
  modnames = samps[["modnames"]], dsids = samps[["dsids"]]
)
```

The scores must be probabilities between 0 and 1. A model that ranks
perfectly but always predicts 0.6 for positives scores badly here, and that
is the point.

## The three metrics

| Metric | What it is | Range | Best |
|---|---|---|---|
| `brier` | Mean squared difference between probability and outcome | 0 to 1 | 0 |
| `rmse` | Square root of the Brier score | 0 to 1 | 0 |
| `logloss` | Mean negative log likelihood | 0 upward | 0 |

The Brier score and the RMSE are the same information on two scales. Log loss
punishes confident mistakes much harder: a probability of 0.01 on something
that turns out positive costs far more than a probability of 0.4 does.

```{r}
knitr::kable(head(prob_metrics(mdat)))
```

One row per model, dataset and metric.

## Confidence intervals

`prob_metrics_ci()` summarizes those per-dataset values into an interval for
each model and metric.

```{r}
knitr::kable(prob_metrics_ci(mdat))
```

Like `auc_ci()`, it takes `alpha` for the level and `dtype` for the normal or
t distribution.

## When to use them

Use them when a downstream decision consumes the probability itself - an
expected-cost calculation, a threshold set from a risk budget, a model whose
output feeds another model. Then calibration matters as much as ranking.

Use ROC and precision-recall curves when the decision is a ranking or a
cutoff. A well-ranked, badly calibrated model is fine for those, and these
metrics will tell you it is bad.

Reporting both is reasonable, and they answer different questions.

## Also available per cutoff

`sar` averages accuracy, the ROC area and one minus the RMSE into a single
per-cutoff measure. It is on the [ranking and
cost](measures-ranking-cost.html) page, and it needs probabilities for the
same reason these do.

## Next

- [AUC and other curve summaries](measures-auc.html)
- [Measures overview](measures-overview.html)
