---
title: "Classification report"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Every other page here reports a metric over all of its cutoffs, or a number
that summarizes the whole curve. This one reports a single operating point,
in the layout `scikit-learn`'s `classification_report` prints: a row per
class, then the rows that summarize them.

```{r}
library(precrec)

data(C3N150)
mdat <- mmdata(C3N150$scores, C3N150$labels)

classification_report(mdat, at = 0.5)
```

## `at` has no default

`scikit-learn` reports on `y_pred`, so by the time the function is called
someone has already decided where to cut. `precrec` holds scores and
evaluates every cutoff, so the report has to be told which one to use, and
the answer is not a detail:

```{r}
strict <- classification_report(mdat, at = 0.7)
strict[strict$class == "c1", ]
```

There is also no threshold that means the same thing on every score scale.
`0.5` is the middle of a probability, but it says nothing about a log-odds
or an SVM margin, and `precrec` accepts all of them. So the choice stays
with you. Pass one number for every class, or one per class:

```{r}
classification_report(mdat, at = c(c1 = 0.4, c2 = 0.5, c3 = 0.6))
```

## Why there is no accuracy row

A multi-class dataset is evaluated by one-vs-rest, and each class is
thresholded on its own. Nothing makes those decisions agree, so an
observation can clear the bar for no class at all, or for several:

```{r}
n_classes <- rowSums(C3N150$scores >= 0.5)
table(n_classes)
```

With observations in no class and in three, there is no single-label
accuracy to report. This is the case `scikit-learn` documents, and it prints
`micro avg` in exactly this situation - precision, recall and F-score
recomputed from the true positives, false positives and false negatives
pooled over the classes, rather than averaged from the per-class rates.

A binary problem is different. One threshold puts every observation on
exactly one side, so accuracy is defined and that is the row you get:

```{r}
data(P10N10)
classification_report(
  scores = P10N10$scores, labels = P10N10$labels,
  at = 12
)
```

Both classes get a row, as they do in `scikit-learn`. The negative row is
the confusion matrix turned around, so its precision is the negative
predictive value and its recall the specificity.

## The three averages

`macro avg` weights every class equally and `weighted avg` weights each by
its support, the same pair [`auc()`](metrics-auc.html) offers as
`macro_weight`. Use the macro average when the rare classes are the
interesting ones, and the weighted average when you want the average
observation.

`micro avg` is neither: it pools the counts first and computes the metrics
once. When the predictions happen to be single-label the three metrics
collapse to the same number, and that number is the accuracy - which is why
`scikit-learn` prints one row or the other and never both.

## Undefined cells

A class nothing was predicted into has no precision, and a class with no
observations has no recall. `zero_division` decides what goes in the cell:
`0` by default, as in `scikit-learn`, or `NA` for the convention the
per-cutoff metrics of [`evalmod()`](metrics-overview.html) follow.

```{r}
mdat_none <- mmdata(c(0, 0, 0, 0), c(1, 1, 0, 0))
classification_report(mdat_none, at = 0.5, zero_division = NA)
```

A score of `NA` is never a positive prediction, matching the `na_worst =
TRUE` default of the rest of the package.

## What it is not

The report is one point on the curves, chosen by you. It cannot tell you
whether a different threshold would have done better, and a threshold picked
by looking at the same data it is scored on will flatter the model. For the
threshold-free view, which is what `precrec` exists for, see
[AUC and other curve summaries](metrics-auc.html).

## Next

- [Metrics overview](metrics-overview.html)
- [AUC and other curve summaries](metrics-auc.html)
