---
title: "Classification report"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Every other page here reports a metric over all of its cutoffs, or a number
that summarizes a whole curve. This one reports a single operating point, in
the layout `scikit-learn`'s `classification_report` prints.

```{r}
library(precrec)

mdat <- mmdata(C3N150$scores, C3N150$labels)

classification_report(mdat, at = 0.5)
```

## `at` has no default

`scikit-learn` reports on `y_pred`, so by the time it is called someone has
already decided where to cut. `precrec` holds scores and evaluates every
cutoff, so the report has to be told which one to use - and the answer is not
a detail. There is also no threshold that means the same thing on every score
scale `precrec` accepts: `0.5` is the middle of a probability but says
nothing about a log-odds or an SVM margin.

Pass one number for every class, or one per class:

```{r}
classification_report(mdat, at = c(c1 = 0.4, c2 = 0.5, c3 = 0.6))
```

## Why there is no accuracy row

A multi-class dataset is evaluated one-vs-rest, and each class is thresholded
on its own. Nothing makes those decisions agree, so an observation can clear
the bar for no class at all, or for several:

```{r}
table(n_classes = rowSums(C3N150$scores >= 0.5))
```

With observations in no class and in three, there is no single-label accuracy
to report. This is the case `scikit-learn` documents, and it prints
`micro avg` in exactly this situation: precision, recall and F-score
recomputed from the true positives, false positives and false negatives
pooled over the classes, rather than averaged from the per-class rates.

A binary problem is different. One threshold puts every observation on
exactly one side, so accuracy is defined and that is the row you get:

```{r}
classification_report(
  scores = P10N10$scores, labels = P10N10$labels,
  at = 12
)
```

Both classes get a row, as in `scikit-learn`. The negative row is the
confusion matrix turned around, so its precision is the negative predictive
value and its recall the specificity.

## The three averages

`macro avg` weights every class equally and `weighted avg` weights each by
its support - the same pair [`auc()`](metrics-auc.html) offers as
`macro_weight`. Use macro when the rare classes are the interesting ones,
weighted when you want the average observation.

`micro avg` is neither: it pools the counts first and computes the metrics
once. When the predictions happen to be single-label the three collapse to
the same number, and that number is the accuracy - which is why
`scikit-learn` prints one row or the other and never both.

## Undefined cells

A class nothing was predicted into has no precision, and a class with no
observations has no recall. `zero_division` decides what goes in the cell:
`0` by default, as in `scikit-learn`, or `NA` for the convention the
per-cutoff metrics follow. A score of `NA` is never a positive prediction,
matching the `na_worst = TRUE` default of the rest of the package.

## What it is not

One point on the curves, chosen by you. It cannot tell you whether a
different threshold would have done better - that is
[`best_cutoff()`](howto-operating-point.html) - and a threshold picked by
looking at the same data it is scored on will flatter the model. For the
threshold-free view, which is what `precrec` exists for, see
[AUC and other curve summaries](metrics-auc.html).

## Next

- [Choose an operating point](howto-operating-point.html)
- [Metrics overview](metrics-overview.html)
- [AUC and other curve summaries](metrics-auc.html)
