---
title: "Uncertainty from one test set"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Every interval elsewhere in `precrec` is built from the variation *between*
test sets, so with one test set there is nothing to vary and `auc_ci()` says
so. `auc_boot()` resamples the one sample you have instead.

```{r}
library(precrec)

samps <- create_sim_samples(1, 100, 100, c("poor_er", "good_er"))
mdat <- mmdata(samps[["scores"]], samps[["labels"]],
  modnames = samps[["modnames"]]
)

booted <- auc_boot(mdat, boot_n = 500, seed = 42)
knitr::kable(auc_ci(booted))
```

`aucs` is the AUC of the original data and `mean` the average over the
resamples; the gap between them is the bootstrap's estimate of bias. The
bounds are quantiles of the resampled values - a percentile interval - so
they are values the AUC actually took, and cannot fall outside `[0, 1]`.
That is why `dtype` is refused here: there is no distribution being assumed.

Pass `seed` to make an interval reproducible. It restores the random stream
afterwards, so a seeded call does not disturb whatever else you are drawing
from.

## Stratified, and shared

Positives are drawn from the positives and negatives from the negatives, so
every resample has exactly the class balance of the original. An
unstratified bootstrap of imbalanced data varies the balance from resample
to resample, and can draw one with no positives at all - which moves the
precision-recall baseline underneath the number being estimated.

Every model is resampled on the *same* draws, which is what makes the next
section a comparison rather than two separate answers.

## Comparing two models

```{r}
knitr::kable(auc_diff(booted))
```

The difference is taken within each resample, so the interval describes the
difference itself. That is not the same as looking at two intervals from
`auc_ci()` and checking whether they overlap: two intervals can overlap
while the difference is clearly on one side of zero, because separate
intervals say nothing about how the two models move together.

`p_values` is the share of resampled differences on the far side of zero
from the observed one, doubled. It is floored at `2 / (n + 1)`, so 500
resamples cannot report anything below 0.004 - a smaller number would be an
artifact of the resample count rather than evidence. Read the interval
first; the p-value is a companion to it, not an exact test.

## The statistic and the second p-value

That floor is a problem when the result has to be written down as a
p-value. Take a pair of models that are not close:

```{r}
clear <- create_sim_samples(1, 100, 100, c("poor_er", "excel"))
cmdat <- mmdata(clear[["scores"]], clear[["labels"]],
  modnames = clear[["modnames"]]
)

clear_diff <- auc_diff(auc_boot(cmdat, boot_n = 500, seed = 42))
# The normal p-value is far below what kable's default rounding shows
clear_diff[["p_values_norm"]] <- format(
  clear_diff[["p_values_norm"]],
  digits = 3
)

knitr::kable(clear_diff)
```

The interval is nowhere near zero, and yet `p_values` sits on its floor of
0.004, because that is the smallest number 500 resamples can support.
Getting below 0.001 that way means resampling ten thousand times or more.

`statistic` and `p_values_norm` read a number off the resamples that are
already there. `statistic` is a Wald statistic - an estimate over an
estimate of its standard error. The estimate is the observed difference,
and the standard error is the standard deviation of the resampled
differences, since the spread of the bootstrap distribution is what the
bootstrap has to say about how far the difference moves from sample to
sample. `p_values_norm` reads that off the standard normal, here many
orders of magnitude below the floor the percentile p-value has run into.

It is not a *t* statistic, and the difference is worth seeing, because
`auc_ci(dtype = "t")` on cross-validation folds is one. There the spread is
taken over a handful of real test sets and divided by the square root of
how many there were, so `n - 1` degrees of freedom mean something. Here the
spread is taken over resamples and divided by nothing: `boot_n` is a
setting rather than a sample size, and a t with `boot_n - 1` degrees of
freedom would be a p-value that shrinks when you resample harder.

The resolution is bought with two assumptions, and it is worth knowing
which. The first is that the statistic and its p-value are the test
counterpart of the *normal* interval, the difference plus and minus so many
standard errors, while `lower_bound` and `upper_bound` are the *percentile*
interval. `p_values_norm` is therefore not the dual of the bounds next to
it, and the two need not agree.

The second is normality itself. Nothing is shuffled between the two models,
so the null is never enforced: the distribution being used is the sampling
one, centered on the observed difference and reflected to sit under zero.
It also takes the resampled differences to be roughly normal, which is
weakest where they are skewed - few positives, or either model near the
ceiling of the precision-recall AUC. The very small p-values it makes
available are therefore its least trustworthy numbers, and when the two
p-values disagree sharply it is the normal approximation to distrust.

`statistic` is `NA` when the resamples have no spread to divide by, which
is what two models given the same scores produce, and `p_values_norm` is
`NA` with it.

## One side or two

`alternative` picks the tail both p-values are read from. The default
`"two.sided"` asks whether the models differ; `"greater"` asks whether the
first of the pair has the larger AUC, and `"less"` whether the second does.
`poor_er` is the first of this pair, so `"less"` is the question worth
asking about it:

```{r}
knitr::kable(auc_diff(booted, alternative = "less"))
```

The interval stays two-sided whatever `alternative` is set to, so the same
bounds are there to read next to a one-sided p-value.

## How many resamples

`boot_n` defaults to 1000. The tails of a 95% interval are estimated from
about 25 resamples at each end of that, which is enough to see the interval
but coarse if the interval is the result you are publishing; 2000 or more is
worth the wait then. Cost grows with `boot_n` and barely with the size of
the dataset, since the curves are computed in C++ and all the resamples go
through in one pass.

## When several test sets are better

If you have cross-validation folds or genuinely separate test sets, use
them: `auc_ci()` on an `evalmod()` result measures variation between real
samples, while the bootstrap can only reuse the one sample it was given. A
bootstrap interval inherits whatever is unrepresentative about that sample,
and cannot tell you so.

## Next

- [AUC and other curve summaries](metrics-auc.html)
- [Evaluate cross-validation folds](howto-cross-validation.html)
