---
title: "Comparison with other tools"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

`precrec` overlaps with several other packages. This page is about which to
reach for, and about the one place where they disagree on a number rather
than on an interface.

| Tool | What it is for | Against `precrec` |
|---|---|---|
| `ROCR` (R) | Cutoff metrics and their plots | `precrec` covers its metrics and accepts its names; see below |
| `pROC` (R) | ROC analysis and inference | More ways to compare two ROC curves, and ROC curve smoothing; no precision-recall curves |
| `PRROC` (R) | Precision-recall curves | Also integrates them properly, and takes weighted or soft labels, which `precrec` does not |
| `yardstick` (R) | Metrics inside `tidymodels` | Tibble in, tibble out, and fits `tune` and `workflows`; `precrec` reached name parity in 0.17.0 |
| `scikit-learn` (Python) | General model evaluation | The reference in Python; `classification_report()` mirrors its report |
| `imbalanced-learn` (Python) | Metrics for imbalanced data | Has the geometric mean and an index-balanced accuracy that `precrec` does not |

## Where the numbers differ

There are two ways to turn precision-recall points into an area, and they
are not two ways of computing the same thing.

The **step** estimator joins the points with horizontal steps - what
`scikit-learn`'s `average_precision_score` reports, and what `precrec`
provides as `average_precision()`. The **interpolated** area is what `auc()`
returns for the PRC, taken along the non-linear interpolation between
achievable points. On balanced data they nearly agree; as the positives get
rarer they do not:

```{r}
library(precrec)

compare <- function(np, nn) {
  set.seed(1)
  scores <- c(rnorm(np, 1.2), rnorm(nn, 0))
  labels <- rep(c(1, 0), c(np, nn))
  curves <- evalmod(scores = scores, labels = labels)
  areas <- auc(curves)

  data.frame(
    positives = sprintf("%.0f%%", 100 * np / (np + nn)),
    interpolated = areas$aucs[areas$curvetypes == "PRC"],
    step = average_precision(curves)$aps
  )
}

res <- do.call(rbind, list(
  compare(500, 500), compare(100, 900),
  compare(20, 980)
))
res$overstated <- sprintf("%+.1f%%", 100 * (res$step / res$interpolated - 1))

knitr::kable(res, digits = 4)
```

At two percent positives the step estimator reads more than a tenth too
high. That is the case `precrec` was written for, and it is why the
interpolated area is what `auc()` reports. Joining the raw points with
straight lines is a third answer again, and the one the package exists to
argue against.

## ROCR names work here

The ROCR metrics were checked against ROCR itself over balanced, imbalanced
and tied data as they were added, and its names are accepted as aliases:

```{r}
points <- evalmod(
  scores = P10N10$scores, labels = P10N10$labels,
  mode = "basic", metrics = c("fall", "miss", "rpp")
)

unique(as.data.frame(points)$type)
```

`fall`, `miss` and `rpp` come back as `fpr`, `fnr` and
`predicted_positive_rate`. Three ROCR metrics are deliberately absent - the
ROC convex hull, the expected-cost curve and the calibration error - each
being a curve in a space of its own rather than a column of the metric
table.

## When to use something else

Reach for `pROC` if you need ROC inference `precrec` does not have:
Venkatraman's test for whole curves, tests at a fixed sensitivity or
specificity, comparisons between models fitted on *different* datasets, or
curve smoothing. DeLong's standard error and test are no longer a reason to
switch - [`auc_delong()`](metrics-uncertainty.html) has them - and neither
is `coords(x, "best")`, which is
[`best_cutoff()`](howto-operating-point.html) here, with the criteria split
into the ones that know the prevalence and the ones that do not.
Reach for
`PRROC` if your labels are weighted or soft. Reach for `yardstick` if the surrounding
code is `tidymodels`, and for `scikit-learn` if it is Python.

Reach for `precrec` when the precision-recall curve is the point, when the
data are imbalanced enough for the table above to matter, or when you want
one call to give you the curves, the per-cutoff metrics and the areas
together.

If you have decided and want the call-by-call mapping rather than the
argument, it is on [Coming from pROC or ROCR](howto-from-proc-rocr.html).

## Next

- [Coming from pROC or ROCR](howto-from-proc-rocr.html)
- [Metrics overview](metrics-overview.html)
- [AUC and other curve summaries](metrics-auc.html)
