---
title: "AUC and other curve summaries"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

These summarize a whole curve in one number, so they do not depend on picking
a cutoff.

```{r}
library(precrec)

curves <- evalmod(scores = P10N10$scores, labels = P10N10$labels)

knitr::kable(auc(curves))
```

| Curve | What the area means |
|---|---|
| ROC | Probability a random positive outranks a random negative |
| PRC | Mean precision across all recall levels |

`precrec` computes the PRC area from the properly interpolated curve; tools
that join raw points with straight lines overestimate it.

## Read the area against its baseline

The `baselines` column is what the area would be by chance. For a ROC curve
that is 0.5 whatever the data, so a ROC AUC can be read on its own. For a
precision-recall curve it is the proportion of positives, so a PRC AUC
cannot. This is the same generator at three class balances:

```{r}
compare <- function(np, nn) {
  set.seed(1)
  scores <- c(rnorm(np, 1.2), rnorm(nn, 0))
  labels <- rep(c(1, 0), c(np, nn))
  areas <- auc(evalmod(scores = scores, labels = labels))

  data.frame(
    positives = sprintf("%.0f%%", 100 * np / (np + nn)),
    roc = areas$aucs[areas$curvetypes == "ROC"],
    prc = areas$aucs[areas$curvetypes == "PRC"],
    prc_baseline = areas$baselines[areas$curvetypes == "PRC"]
  )
}

knitr::kable(
  do.call(rbind, list(compare(500, 500), compare(100, 900), compare(20, 980))),
  row.names = FALSE, digits = 3
)
```

The classifier is about as good in all three rows and the ROC AUC says so,
holding between 0.81 and 0.85. The PRC AUC falls to 0.129, which reads as
failure and is in fact six times chance. Quote the two numbers together, and
see
[Balanced and imbalanced data](howto-imbalanced-data.html) for why the PRC
column is the one that moved.

The baseline is taken per model and per test dataset, because a fold need not
hold the classes in the proportions the whole dataset does.

## Average precision

A different estimator of the same curve, not a different way of adding up the
same one: it joins the raw points with horizontal steps instead of
interpolating, so it reads high wherever the two disagree - here by about
0.006. Prefer the interpolated area; this is here because other packages
report it under this name.

```{r}
knitr::kable(average_precision(curves))
```

## Partial areas

When only part of the curve matters - the low-false-positive end, say -
`part()` restricts the range and `pauc()` reports the area over it.
`spaucs` are standardized partial AUCs, rescaled to 0 to 1 so ranges of
different widths compare.

```{r}
knitr::kable(pauc(part(curves, xlim = c(0, 0.25))))
```

There is a second convention for standardizing a partial area, and both
are called "the standardized partial AUC". `spaucs` divides by the area of
the region, so it says what fraction of what was available the curve
covered. The McClish correction, which `corrected = TRUE` adds as `cpaucs`,
rescales the span between chance and perfect onto 0.5 to 1 instead, so the
number reads on the same scale as a full AUC - and over the whole curve it
is the full AUC.

```{r}
knitr::kable(
  subset(
    pauc(part(curves, xlim = c(0, 0.25)), corrected = TRUE),
    curvetypes == "ROC"
  ),
  row.names = FALSE
)
```

Neither is the right one; say which you used. `cpaucs` is what
`pROC::auc(..., partial.auc.correct = TRUE)` reports, and it is `NA` on the
precision-recall rows - a precision-recall curve's chance level is the
proportion of positives, and rescaling by it would make curves at different
class balances look comparable, which is the thing this package spends
[a whole page](howto-imbalanced-data.html) arguing they are not. See
`?pauc`.

Plotting the restricted object draws the partial curve; see
[partial curves](plots-partial-curves.html).

## Confidence intervals

With several test sets, `auc_ci()` puts an interval around the mean area.
`alpha` sets the level, and `dtype = "t"` uses the t distribution, the better
choice for a small number of folds.

```{r}
samps <- create_sim_samples(10, 100, 100, "good_er")
mdat <- mmdata(samps[["scores"]], samps[["labels"]], dsids = samps[["dsids"]])
mcurves <- evalmod(mdat)

knitr::kable(auc_ci(mcurves, alpha = 0.01, dtype = "t"))
```

## Break-even point

Where precision and recall are equal - the one point on the curve where the
two are in balance. `prbe()` reads the individual curves, so ask `evalmod()`
to keep them.

```{r}
knitr::kable(prbe(evalmod(mdat, raw_curves = TRUE)))
```

A curve crossing the diagonal more than once gets one row per crossing; one
that never does gets a single `NA`. `prbe()` reads the crossing off the
interpolated curve - finding it by linear interpolation between raw points,
as some tools do, is the error this package exists to avoid.

## Averaging over classes

For a multiclass evaluation `auc()` adds the macro-average of the per-class
areas, weighting every class equally. `macro_weight = "prevalence"` weights
each by the number of observations it has instead, and names those rows
`macro-average-weighted` to keep the two apart.

```{r}
mccurves <- evalmod(scores = C3N150$scores, labels = C3N150$labels)

knitr::kable(auc(mccurves, macro_weight = "prevalence"))
```

Uniform if every class matters equally, prevalence if you care about the
average case. Other packages call these `roc_aunu` and `roc_aunp`. See
[more than two classes](howto-multiclass.html).

The baseline is averaged over the classes the same way the areas are, with
the same weights, so a macro row stays readable. That matters most when the
classes are the unequal sizes a macro-average is usually reached for:

```{r}
set.seed(5)
rare <- rep(c("c1", "c2", "c3"), c(10, 40, 150))
noise <- cbind(rnorm(200), rnorm(200), rnorm(200))
noisy <- evalmod(mmdata(noise, rare, multiclass = "ovr"))

knitr::kable(
  subset(
    auc(noisy, macro_weight = "prevalence"),
    curvetypes == "PRC"
  ),
  row.names = FALSE, digits = 3
)
```

The scores here are random numbers. The weighted macro-average PRC area is
0.584, which would be a respectable result if the baseline were not 0.605
sitting beside it.

## The fast path

If the ROC area is all you need, `mode = "aucroc"` computes it from the U
statistic of the Mann-Whitney test without building the curve. See
[large datasets](howto-large-datasets.html).

```{r}
knitr::kable(as.data.frame(
  evalmod(scores = P10N10$scores, labels = P10N10$labels, mode = "aucroc")
))
```

## Next

- [Uncertainty from one test set](metrics-uncertainty.html)
- [Probability-based metrics](metrics-probability.html)
