---
title: "Uncertainty from one test set"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Every interval elsewhere in `precrec` is built from the variation *between*
test sets, so with one test set there is nothing to vary and `auc_ci()` says
so. `auc_boot()` resamples the one sample you have instead.

```{r}
library(precrec)

samps <- create_sim_samples(1, 100, 100, c("poor_er", "good_er"))
mdat <- mmdata(samps[["scores"]], samps[["labels"]],
  modnames = samps[["modnames"]]
)

booted <- auc_boot(mdat, boot_n = 500, seed = 42)
knitr::kable(auc_ci(booted))
```

`aucs` is the AUC of the original data and `mean` the average over the
resamples; the gap between them is the bootstrap's estimate of bias. The
bounds are quantiles of the resampled values - a percentile interval - so
they are values the AUC actually took, and cannot fall outside `[0, 1]`.
That is why `dtype` is refused here: there is no distribution being assumed.

Pass `seed` to make an interval reproducible. It restores the random stream
afterwards, so a seeded call does not disturb whatever else you are drawing
from.

## Stratified, and shared

Positives are drawn from the positives and negatives from the negatives, so
every resample has exactly the class balance of the original. An
unstratified bootstrap of imbalanced data varies the balance from resample
to resample, and can draw one with no positives at all - which moves the
precision-recall baseline underneath the number being estimated.

Every model is resampled on the *same* draws, which is what makes the next
section a comparison rather than two separate answers.

## Comparing two models

```{r}
knitr::kable(auc_diff(booted))
```

The difference is taken within each resample, so the interval describes the
difference itself. That is not the same as looking at two intervals from
`auc_ci()` and checking whether they overlap: two intervals can overlap
while the difference is clearly on one side of zero, because separate
intervals say nothing about how the two models move together.

`p_values` is the *percentile* p-value. Writing `d` for the resampled
differences and `n` for how many there are, it is the share of `d` on the
far side of zero from the observed difference, doubled:

```
p_values = 2 * min(1 + sum(d <= 0), 1 + sum(d >= 0)) / (n + 1)
```

The added ones keep it away from exactly zero, which floors it at
`2 / (n + 1)`: 500 resamples cannot report anything below 0.004, and a
smaller number would be an artifact of the resample count rather than
evidence. Read the interval first; the p-value is a companion to it, not an
exact test.

## The Wald test

That floor is a problem when the result has to be written down as a
p-value. Take a pair of models that are not close:

```{r}
clear <- create_sim_samples(1, 100, 100, c("poor_er", "excel"))
cmdat <- mmdata(clear[["scores"]], clear[["labels"]],
  modnames = clear[["modnames"]]
)

clear_diff <- auc_diff(auc_boot(cmdat, boot_n = 500, seed = 42))
# The Wald p-value is far below what kable's default rounding shows
clear_diff[["p_values_wald"]] <- format(
  clear_diff[["p_values_wald"]],
  digits = 3
)

knitr::kable(clear_diff)
```

The interval is nowhere near zero, and yet `p_values` sits on its floor of
0.004, because that is the smallest number 500 resamples can support.
Getting below 0.001 that way means resampling ten thousand times or more.

`z_values` and `p_values_wald` are a *Wald* test, read off the resamples
that are already there. A Wald statistic is an estimate over an estimate of
its standard error, referred to a standard normal:

```
z_values      = diffs / sd(d)
p_values_wald = 2 * pnorm(-abs(z_values))
```

`sd(d)` is the standard error, because the spread of the bootstrap
distribution is what the bootstrap has to say about how far the difference
moves from sample to sample. Having a scale underneath it rather than a
count of resamples, it has no floor, and lands many orders of magnitude
below the one the percentile p-value has run into.

`z_values` is `NA` when the resamples have no spread to divide by, which is
what two models given the same scores produce, and `p_values_wald` is `NA`
with it.

## Why two p-values

Because they fail in opposite ways, and neither one on its own tells you
that it is failing.

The percentile p-value assumes nothing about the shape of `d`, and pays for
that with the floor. Once it reaches `2 / (n + 1)` it has stopped measuring
the models and started reporting `boot_n` - a difference that is merely
clear and one that is overwhelming both come out at 0.004 in the table
above, and nothing in the number says which one you are looking at.

The Wald p-value has no floor, and pays for that with an assumption. It
takes `d` to be roughly normal, and locates the null by reflecting the
sampling distribution rather than by enforcing it: nothing is shuffled
between the two models. Where `d` is skewed - few positives, or either
model near the ceiling of the precision-recall AUC - it is confidently
wrong, and again the number carries no warning.

Side by side they cover each other. When they agree, the normality the Wald
test assumes is doing no harm at this sample size, and its resolution is
yours to quote. When they disagree sharply, `d` is not the shape the Wald
test needs, and the percentile p-value - floor and all - is the one to
trust. The comparison is the diagnostic; neither column is one by itself.

There is also an exact sense in which they answer different questions. A
Wald test is the counterpart of the *normal* interval, the difference plus
and minus so many standard errors, while `lower_bound` and `upper_bound`
are the *percentile* interval. `p_values_wald` is therefore not the dual of
the bounds next to it, and need not agree with them.

## Not a t statistic

`z_values` divides by a standard error, not by a standard error of a mean,
and is read off the normal rather than off a t. The contrast with
`auc_ci(dtype = "t")` on cross-validation folds, which is a genuine t, is
exact: there the spread is taken over a handful of real test sets and
divided by the square root of how many there were, so `n - 1` degrees of
freedom mean something. Here the spread is taken over resamples and divided
by nothing - `boot_n` is a setting rather than a sample size, and a t on
`boot_n - 1` degrees of freedom would be a p-value that shrinks when you
resample harder.

Nor is it the bootstrap-*t*, which forms a statistic of its own inside
every resample and takes its reference distribution from those rather than
from the normal.

## One side or two

`alternative` picks the tail the percentile and the Wald p-value are read
from. The default
`"two.sided"` asks whether the models differ; `"greater"` asks whether the
first of the pair has the larger AUC, and `"less"` whether the second does.
`poor_er` is the first of this pair, so `"less"` is the question worth
asking about it:

```{r}
knitr::kable(auc_diff(booted, alternative = "less"))
```

The interval stays two-sided whatever `alternative` is set to, so the same
bounds are there to read next to a one-sided p-value.

## How many resamples

`boot_n` defaults to 1000. The tails of a 95% interval are estimated from
about 25 resamples at each end of that, which is enough to see the interval
but coarse if the interval is the result you are publishing; 2000 or more is
worth the wait then. Cost grows with `boot_n` and barely with the size of
the dataset, since the curves are computed in C++ and all the resamples go
through in one pass.

## When several test sets are better

If you have cross-validation folds or genuinely separate test sets, use
them: `auc_ci()` on an `evalmod()` result measures variation between real
samples, while the bootstrap can only reuse the one sample it was given. A
bootstrap interval inherits whatever is unrepresentative about that sample,
and cannot tell you so.

## Next

- [AUC and other curve summaries](metrics-auc.html)
- [Evaluate cross-validation folds](howto-cross-validation.html)
