---
title: "Metrics overview"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

`evalmod(mode = "basic")` evaluates metrics at every cutoff of the score,
rather than at one arbitrary threshold. This page is the map; each family has
its own short page.

```{r}
library(precrec)
library(ggplot2)

points <- evalmod(scores = P10N10$scores, labels = P10N10$labels,
  mode = "basic"
)
```

## The default fourteen

These are calculated whenever you ask for `mode = "basic"`, and each gets a
panel in the default plot.

| Metric | Short | Page |
|---|---|---|
| `score` | `score` | The score itself, for reference |
| `label` | `label` | The observed label, for reference |
| `error` | `err` | [Confusion-matrix rates](metrics-confusion-matrix.html) |
| `accuracy` | `acc` | [Confusion-matrix rates](metrics-confusion-matrix.html) |
| `specificity` | `sp` | [Confusion-matrix rates](metrics-confusion-matrix.html) |
| `sensitivity` | `sn` | [Confusion-matrix rates](metrics-confusion-matrix.html) |
| `precision` | `prec` | [Confusion-matrix rates](metrics-confusion-matrix.html) |
| `npv` | `npv` | [Confusion-matrix rates](metrics-confusion-matrix.html) |
| `balanced_accuracy` | `bacc` | [Agreement and balance](metrics-agreement.html) |
| `fscore` | `fscore` | [Agreement and balance](metrics-agreement.html) |
| `mcc` | `mcc` | [Agreement and balance](metrics-agreement.html) |
| `kappa` | `kappa` | [Agreement and balance](metrics-agreement.html) |
| `informedness` | `infm` | [Agreement and balance](metrics-agreement.html) |
| `markedness` | `mkd` | [Agreement and balance](metrics-agreement.html) |

## The seventeen you ask for

Each of these is another vector the size of your dataset and another panel in
the plot, so they are left out unless named.

| Metric | Also known as | Page |
|---|---|---|
| `fpr` | `fall` | [Confusion-matrix rates](metrics-confusion-matrix.html) |
| `fnr` | `miss` | [Confusion-matrix rates](metrics-confusion-matrix.html) |
| `false_discovery_rate` | `fdr`, `pcfall` | [Confusion-matrix rates](metrics-confusion-matrix.html) |
| `false_omission_rate` | `for`, `pcmiss` | [Confusion-matrix rates](metrics-confusion-matrix.html) |
| `predicted_positive_rate` | `ppr`, `rpp` | [Confusion-matrix rates](metrics-confusion-matrix.html) |
| `predicted_negative_rate` | `pnr`, `rnp` | [Confusion-matrix rates](metrics-confusion-matrix.html) |
| `lift` | | [Ranking and cost](metrics-ranking-cost.html) |
| `odds` | `odds_ratio` | [Ranking and cost](metrics-ranking-cost.html) |
| `positive_likelihood_ratio` | `lrp` | [Ranking and cost](metrics-ranking-cost.html) |
| `negative_likelihood_ratio` | `lrn` | [Ranking and cost](metrics-ranking-cost.html) |
| `mi` | `mutual_information` | [Ranking and cost](metrics-ranking-cost.html) |
| `chisq` | | [Ranking and cost](metrics-ranking-cost.html) |
| `cost` | | [Ranking and cost](metrics-ranking-cost.html) |
| `sar` | | [Ranking and cost](metrics-ranking-cost.html) |
| `roc_dist` | | [Agreement and balance](metrics-agreement.html) |
| `sedi` | | [Agreement and balance](metrics-agreement.html) |
| `jaccard` | `jacc` | [Confusion-matrix rates](metrics-confusion-matrix.html) |

## Asking for them

`metrics =` names the extra metrics. The default fourteen are always kept,
so nothing that worked before changes.

```{r}
extra <- evalmod(scores = P10N10$scores, labels = P10N10$labels,
  mode = "basic", metrics = c("fpr", "lift")
)

autoplot(extra, c("fpr", "lift"))
```

`metrics = "all"` asks for every metric at once.

Any of the names in the tables above works wherever a metric is named -
in `metrics =`, in `plot()` and `autoplot()`, and on either axis of
[`metric_curve()`](plots-metric-curve.html). Names given to these metrics by
other tools are accepted too, so a call written against `ROCR` keeps working.

## Not on this page

Three summaries describe a whole curve or a whole set of predictions rather
than one cutoff, and have their own functions:

- [AUC and other curve summaries](metrics-auc.html) - `auc()`, `pauc()`,
  `auc_ci()`, `prbe()`, `average_precision()`
- [Probability-based metrics](metrics-probability.html) - `prob_metrics()`,
  for scores that are genuine probabilities, and the D2 scores that rescale
  its losses against a null model
