---
title: "Ranking and cost"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Metrics that answer questions the confusion-matrix rates do not: how much
better than random is this cutoff, how much does it cost, how much does it
tell us. All of them are opt-in through `metrics =`.

```{r}
library(precrec)
library(ggplot2)

samps <- create_sim_samples(1, 100, 100, "good_er")
```

## Lift and odds

| Metric | Formula | Range |
|---|---|---|
| `lift` | sensitivity / rate of positive predictions | 0 upward |
| `odds` | (TP x TN) / (FN x FP) | 0 upward |

Lift is how many times better than random selection the cutoff is: a lift of
3 means the flagged group holds three times the share of positives the whole
dataset does. It is the standard metric in marketing and screening, where the
question is what to do with a limited review budget. The odds ratio is the
odds of being positive among the flagged against the odds among the rest.

```{r}
points <- evalmod(
  scores = samps$scores, labels = samps$labels,
  mode = "basic", metrics = c("lift", "odds")
)

autoplot(points, "lift")
```

## Likelihood ratios

| Metric | Formula | Range |
|---|---|---|
| `positive_likelihood_ratio` | sensitivity / FPR | 0 upward |
| `negative_likelihood_ratio` | FNR / specificity | 0 upward |

The two halves of the odds ratio - LR+ divided by LR- is exactly `odds` -
kept apart because they answer different questions. LR+ is how much a
positive prediction multiplies the odds that a case really is positive, LR-
how much a negative one does. A test can be worth using on one alone, and the
odds ratio, being the quotient, hides which is doing the work.

Both are ratios of rates rather than counts, so neither moves with the
prevalence. That is what lets a value carry from one population to another,
and also the catch: a cutoff with an excellent LR+ still flags mostly false
positives if positives are rare enough. Precision, on the
[confusion-matrix page](metrics-confusion-matrix.html), answers *that*.

## Cost

`cost` weights the two kinds of error separately, as
`(cost_fp x FP + cost_fn x FN) / all`. With the default weights of 1 it is
the error rate. Set them to what the two mistakes actually cost you, and the
minimum of the curve is the cutoff to use.

```{r}
costs <- evalmod(
  scores = samps$scores, labels = samps$labels,
  mode = "basic", metrics = "cost", cost_fp = 3, cost_fn = 0.5
)

autoplot(costs, "cost")
```

The metric is not normalized, following `ROCR`, so its scale is the scale of
the weights you gave.

## Information

| Metric | What it is | Range |
|---|---|---|
| `mi` | Mutual information between prediction and truth, in bits | 0 to 1 |
| `chisq` | Pearson chi-square of the 2x2 table, `n x mcc^2` | 0 upward |

Both ask how far the table is from independence rather than how good the
predictions are. They are symmetric: a perfectly wrong classifier scores as
high as a perfectly right one.

## SAR

`sar` is the mean of accuracy, the ROC area, and one minus the root mean
squared error - a single summary harder to game than any one of them. The
RMSE part reads score values rather than ranks, so it needs probabilities
between 0 and 1; given anything else it warns and returns `NA`, and every
other metric in the same call is still returned.

## Where values are undefined

Several of these are undefined at the top and bottom of the ranking, where
the 2x2 table has an empty cell.

- `odds` and `chisq` are `NA` there. Some tools report an infinity or a
  `NaN`; `precrec` reports `NA`, as it does for the undefined end of
  precision and NPV.
- The likelihood ratios are `NA` over a longer stretch, not only at the ends.
  LR+ divides by the false positive rate, 0 for every cutoff above the
  highest-scoring negative; LR- divides by the specificity, 0 from the point
  where every negative has been flagged onward.
- `mi` is `0` there rather than `NA`. A cutoff that predicts one class for
  everything carries no information about the labels, so the value is defined
  and it is zero.

## Next

- [AUC and other curve summaries](metrics-auc.html)
- [One metric against another](plots-metric-curve.html)
