---
title: "Confusion-matrix rates"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Everything here is a ratio of two cells of the 2x2 table, evaluated at every
cutoff.

|  | Predicted positive | Predicted negative |
|---|---|---|
| **Actually positive** | TP | FN |
| **Actually negative** | FP | TN |

```{r}
library(precrec)
library(ggplot2)

points <- evalmod(
  scores = P10N10$scores, labels = P10N10$labels,
  mode = "basic"
)
```

## Overall

| Metric | Formula | Range |
|---|---|---|
| `accuracy` | (TP + TN) / all | 0 to 1 |
| `error` | (FP + FN) / all | 0 to 1 |

Accuracy is the metric to distrust first. On data with 1% positives, calling
everything negative scores 0.99.

```{r}
autoplot(points, c("accuracy", "error"))
```

## Rates over the actual classes

These divide by a row, so they do not move when the class balance changes.

| Metric | Formula | Also known as |
|---|---|---|
| `sensitivity` | TP / (TP + FN) | recall, TPR, hit rate |
| `specificity` | TN / (TN + FP) | TNR, selectivity |
| `fpr` | FP / (FP + TN) | fall-out, 1 - specificity |
| `fnr` | FN / (TP + FN) | miss rate, 1 - sensitivity |

The ROC curve is `sensitivity` against `fpr`, which is why it is blind to
class balance - both axes are row-wise rates.

## Rates over the predicted classes

These divide by a column, so they *do* move with the class balance - what
makes them informative on imbalanced data, and impossible to transfer
between datasets.

| Metric | Formula | Also known as |
|---|---|---|
| `precision` | TP / (TP + FP) | PPV |
| `npv` | TN / (TN + FN) | negative predictive value |
| `false_discovery_rate` | FP / (TP + FP) | 1 - precision |
| `false_omission_rate` | FN / (TN + FN) | 1 - NPV |

```{r}
autoplot(points, c("sensitivity", "precision"))
```

The precision-recall curve is `precision` against `sensitivity` - one
column-wise rate against one row-wise rate. See
[imbalanced data](howto-imbalanced-data.html).

## How much gets flagged

| Metric | Formula | Also known as |
|---|---|---|
| `predicted_positive_rate` | (TP + FP) / all | rate of positive predictions |
| `predicted_negative_rate` | (TN + FN) / all | rate of negative predictions |

Useful when review cost is the constraint: they say how much work a cutoff
creates, regardless of whether the work is well spent.

## Leaving out the true negatives

| Metric | Formula | Also known as |
|---|---|---|
| `jaccard` | TP / (TP + FP + FN) | critical success index, threat score |

The 2x2 table with one corner deleted. The true negatives do not count, the
same omission `precision` and `sensitivity` make, which is why the three move
together on imbalanced data while accuracy does not.

Pad the dataset with 200 negatives that score below everything real, and
nothing about its handling of the positives has changed:

```{r}
best <- function(s, l, m) {
  df <- as.data.frame(evalmod(
    scores = s, labels = l, mode = "basic",
    metrics = m
  ))
  max(df$y[df$type == m], na.rm = TRUE)
}

pad_s <- c(P10N10$scores, rep(0, 200))
pad_l <- c(P10N10$labels, rep(-1, 200))

c(
  jaccard = best(pad_s, pad_l, "jaccard"),
  accuracy = best(pad_s, pad_l, "accuracy")
)
```

`jaccard` holds at 0.5625; accuracy climbs from 0.70, because most of what it
now counts is the padding. Forecast verification calls this the critical
success index for the same reason: a rare event has so many true negatives
that any metric counting them reports mostly the rarity.

## Undefined ends

Precision has no value where nothing is predicted positive, and NPV none
where nothing is predicted negative; `precrec` fills each in from its
neighbor rather than dropping the point. `jaccard` needs no such treatment -
its denominator is 0 only for a dataset with no positives at all.

## Next

- [Agreement and balance](metrics-agreement.html)
- [Ranking and cost](metrics-ranking-cost.html)
