---
title: "AUC and other curve summaries"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

These summarize a whole curve in one number, so they do not depend on picking
a cutoff.

```{r}
library(precrec)

curves <- evalmod(scores = P10N10$scores, labels = P10N10$labels)

knitr::kable(auc(curves))
```

| Curve | What the area means | Baseline |
|---|---|---|
| ROC | Probability a random positive outranks a random negative | 0.5 |
| PRC | Mean precision across all recall levels | Proportion of positives |

The ROC baseline is always 0.5. The precision-recall baseline moves with the
class balance, so a PRC area of 0.4 is poor on balanced data and good when
positives are 5% of the total. `precrec` computes that area from the properly
interpolated curve; tools that join raw points with straight lines
overestimate it.

## Average precision

A different estimator of the same curve, not a different way of adding up the
same one: it joins the raw points with horizontal steps instead of
interpolating, so it reads high wherever the two disagree - here by about
0.006. Prefer the interpolated area; this is here because other packages
report it under this name.

```{r}
knitr::kable(average_precision(curves))
```

## Partial areas

When only part of the curve matters - the low-false-positive end, say -
`part()` restricts the range and `pauc()` reports the area over it.
`spaucs` are standardized partial AUCs, rescaled to 0 to 1 so ranges of
different widths compare.

```{r}
knitr::kable(pauc(part(curves, xlim = c(0, 0.25))))
```

Plotting the restricted object draws the partial curve; see
[partial curves](plots-partial-curves.html).

## Confidence intervals

With several test sets, `auc_ci()` puts an interval around the mean area.
`alpha` sets the level, and `dtype = "t"` uses the t distribution, the better
choice for a small number of folds.

```{r}
samps <- create_sim_samples(10, 100, 100, "good_er")
mdat <- mmdata(samps[["scores"]], samps[["labels"]], dsids = samps[["dsids"]])
mcurves <- evalmod(mdat)

knitr::kable(auc_ci(mcurves, alpha = 0.01, dtype = "t"))
```

## Break-even point

Where precision and recall are equal - the one point on the curve where the
two are in balance. `prbe()` reads the individual curves, so ask `evalmod()`
to keep them.

```{r}
knitr::kable(prbe(evalmod(mdat, raw_curves = TRUE)))
```

A curve crossing the diagonal more than once gets one row per crossing; one
that never does gets a single `NA`. `prbe()` reads the crossing off the
interpolated curve - finding it by linear interpolation between raw points,
as some tools do, is the error this package exists to avoid.

## Averaging over classes

For a multiclass evaluation `auc()` adds the macro-average of the per-class
areas, weighting every class equally. `macro_weight = "prevalence"` weights
each by the number of observations it has instead, and names those rows
`macro-average-weighted` to keep the two apart.

```{r}
mccurves <- evalmod(scores = C3N150$scores, labels = C3N150$labels)

knitr::kable(auc(mccurves, macro_weight = "prevalence"))
```

Uniform if every class matters equally, prevalence if you care about the
average case. Other packages call these `roc_aunu` and `roc_aunp`. See
[more than two classes](howto-multiclass.html).

## The fast path

If the ROC area is all you need, `mode = "aucroc"` computes it from the U
statistic of the Mann-Whitney test without building the curve. See
[large datasets](howto-large-datasets.html).

```{r}
knitr::kable(as.data.frame(
  evalmod(scores = P10N10$scores, labels = P10N10$labels, mode = "aucroc")
))
```

## Next

- [Uncertainty from one test set](metrics-uncertainty.html)
- [Probability-based metrics](metrics-probability.html)
