---
title: "Customizing plots"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

`autoplot()` returns an ordinary `ggplot` object, so the whole of `ggplot2`
applies to it.

```{r}
library(precrec)
library(ggplot2)

curves <- evalmod(scores = P10N10$scores, labels = P10N10$labels)
```

## Add layers to the result

```{r}
autoplot(curves, "PRC") +
  labs(title = "Precision-recall", subtitle = "P10N10") +
  theme_minimal()
```

## Draw the baseline yourself

The plots leave the precision-recall baseline out, because it sits at the
proportion of positives and that differs between datasets. When you are
looking at one dataset you know the number, so add it.

```{r}
baseline <- sum(P10N10$labels == 1) / length(P10N10$labels)

autoplot(curves, "PRC") +
  geom_hline(yintercept = baseline, linetype = "dashed", color = "grey40")
```

## Start from the data instead

`fortify()` is the `ggplot2` hook, so a `precrec` object can go straight into
`ggplot()` when you want to build the plot from scratch.

```{r}
df <- fortify(curves)
head(df)

ggplot(subset(df, curvetype == "PRC"), aes(x = x, y = y)) +
  geom_line(color = "steelblue", linewidth = 1) +
  coord_fixed() +
  labs(x = "Recall", y = "Precision") +
  theme_bw()
```

The columns are `x`, `y`, `modname`, `dsid`, `dsid_modname` and `curvetype` -
already in long form, so mapping color or facets to a model is direct.

```{r}
samps <- create_sim_samples(1, 100, 100, "all")
mdat <- mmdata(samps[["scores"]], samps[["labels"]],
  modnames = samps[["modnames"]]
)
mdf <- fortify(evalmod(mdat))

ggplot(subset(mdf, curvetype == "PRC"), aes(x = x, y = y, color = modname)) +
  geom_line() +
  facet_wrap(~modname) +
  theme_bw() +
  theme(legend.position = "none")
```

## Getting the grob

Multi-panel output - both curve types at once, or several measures - is
assembled from separate plots. `ret_grob = TRUE` returns that assembled
object instead of drawing it, for placing in a larger layout.

```{r, eval = FALSE}
g <- autoplot(curves, ret_grob = TRUE)
```

## Base graphics

`plot()` takes the usual base arguments, so titles, colors and line types
work the way they do everywhere else.

```{r}
plot(curves, "PRC")
```

## Next

- [Plots overview](plots-overview.html)
- [Get the numbers out](howto-results-as-data.html)
