---
title: "One metric against another"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

`metric_curve()` takes the name of a metric for the x axis and the name of a
metric for the y axis and draws one against the other. Every metric
`evalmod()` can calculate is available on both axes.

```{r}
library(precrec)
library(ggplot2)

samps <- create_sim_samples(1, 100, 100, "good_er")
```

## The familiar pair

```{r}
xy <- metric_curve(
  scores = samps[["scores"]], labels = samps[["labels"]],
  x_metric = "fpr", y_metric = "sensitivity"
)

autoplot(xy)
```

That pair is the ROC curve, and it is the default.

## Anything else, as points

```{r}
xy2 <- metric_curve(
  scores = samps[["scores"]], labels = samps[["labels"]],
  x_metric = "predicted_positive_rate", y_metric = "lift"
)

autoplot(xy2)
```

Note the points. That is deliberate, and it is the one thing worth
understanding about this function.

## Which pairs are joined by a line

`precrec` exists because the points of a precision-recall curve must not be
joined by straight lines. The metrics this function reads are raw per-cutoff
values with no interpolation, so joining an arbitrary pair of them would be
the very error the package was written to avoid.

Two pairs have a defined interpolation, and only those two are drawn as
curves:

| x | y | Curve |
|---|---|---|
| `fpr` | `sensitivity` | ROC |
| `sensitivity` | `precision` | Precision-recall |

For those two, `metric_curve()` hands the work to the same code
`evalmod(mode = "rocprc")` uses, so the two cannot disagree.

Everything else is drawn as points. Pass `type = "l"` to join them anyway,
having decided that the straight lines mean something for the pair at hand.

## Several models and test sets

`metric_curve()` draws one curve per test dataset and does not average over
them. An average needs a rule for interpolating between the points of each
curve, which is exactly what an unregistered pair does not have. Use
`evalmod()` for averaged ROC and precision-recall curves.

```{r}
samps2 <- create_sim_samples(3, 100, 100, c("poor_er", "good_er"))
mdat <- mmdata(samps2[["scores"]], samps2[["labels"]],
  modnames = samps2[["modnames"]], dsids = samps2[["dsids"]]
)

xy3 <- metric_curve(mdat, x_metric = "score", y_metric = "precision")

autoplot(xy3)
```

## Naming the metrics

Both axes accept the long name, the short name, and the name other tools use
- `fall` for `fpr`, `rpp` for `predicted_positive_rate`, and so on. The
[metrics overview](metrics-overview.html) lists them all.

## Next

- [Basic metric plots](plots-basic-metrics.html)
- [Customizing plots](plots-customizing.html)
