---
title: "Basic metric plots"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

`mode = "basic"` plots evaluation metrics against the normalized rank of the
scores - that is, against how far down the ranked list the cutoff sits.

```{r}
library(precrec)
library(ggplot2)

points <- evalmod(scores = P10N10$scores, labels = P10N10$labels,
  mode = "basic"
)
```

## Pick the panels

Name the metrics you want. Each becomes a panel.

```{r}
autoplot(points, c("sensitivity", "specificity", "precision"))
```

```{r}
autoplot(points, c("mcc", "fscore"))
```

Called with no metrics, you get all fourteen default panels at once, which
is useful for a first look and too dense for a report.

## Reading the x axis

The x axis runs from 0 to 1 and is the fraction of the dataset above the
cutoff. At x = 0 nothing is predicted positive; at x = 1 everything is. So
the left edge is the strictest cutoff and the right edge the most permissive.

This is what makes the panels comparable across datasets of different sizes.

The value plotted is the *normalized rank*. There is one cutoff above every
instance and one below all of them, so `n` instances give `n + 1` points,
and the cutoff with `k` instances above it sits at `k / n`.

```{r}
n <- 5 # five instances, so six cutoffs

(0:n) / n
```

Counting those points from 1 rather than from 0, the same thing is
`(rank - 1) / (points - 1)`. Ranks count down from the best score, so the
highest-scoring instance is rank 1 and sits at the left edge - the opposite
direction to R's `rank()`, which counts up from the lowest score.

`ties_method` in `evalmod()` sets the rank that tied instances receive, but
the cutoffs still step one instance at a time, so a run of tied scores
spreads across consecutive x values rather than sharing one.

## Scores and labels

Two extra panels show the data behind the metrics rather than a metric: the
score at each rank, and the observed label.

```{r}
autoplot(points, c("score", "label"))
```

The label panel is the quickest way to see whether the positives really are
concentrated at the top of the ranking.

## Extra metrics

Anything beyond the default fourteen is requested with `metrics =` and then
plotted the same way.

```{r}
extra <- evalmod(scores = P10N10$scores, labels = P10N10$labels,
  mode = "basic", metrics = c("fpr", "lift")
)

autoplot(extra, c("fpr", "lift"))
```

Asking to plot a metric that was not calculated is an error that names the
argument to add. See the [metrics overview](metrics-overview.html).

## Axis ranges

Metrics that can go negative - `mcc`, `kappa`, `informedness`, `markedness`
and `label` - are drawn on a -1 to 1 axis. The rest use 0 to 1, and the few
with no natural bound (`lift`, `odds`, `chisq`, `cost`, `score`) are scaled
to the data.

## Next

- [One metric against another](plots-metric-curve.html)
- [Metrics overview](metrics-overview.html)
