---
title: "Partial curves"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Often only part of a curve matters - the low-false-positive end when every
alert costs a review, or the high-recall end when a miss is expensive.
`part()` restricts a curve to that range.

```{r}
library(precrec)
library(ggplot2)

curves <- evalmod(scores = P10N10$scores, labels = P10N10$labels)
```

## Restricting the range

```{r}
partial <- part(curves, xlim = c(0, 0.25))

autoplot(partial)
```

The full curve stays visible in outline and the selected part is highlighted,
so the region is read in context rather than in isolation.

`ylim` restricts the other axis, and both can be given at once.

```{r}
partial_y <- part(curves, ylim = c(0.5, 1))

autoplot(partial_y)
```

## The areas

`pauc()` returns the area over the restricted range.

```{r}
knitr::kable(pauc(partial))
```

`paucs` is the raw area, which is small simply because the range is narrow.
`spaucs` is the standardized version, rescaled to 0 to 1 so that ranges of
different widths can be compared with each other and against a full AUC.

Report the standardized one unless you have a reason not to. A partial AUC of
0.18 sounds poor and may be excellent for the quarter of the axis it covers.

## Several models

`part()` works on any `evalmod()` object, so the comparison carries over.

```{r}
samps <- create_sim_samples(1, 100, 100, "all")
mdat <- mmdata(samps[["scores"]], samps[["labels"]],
  modnames = samps[["modnames"]]
)

mpartial <- part(evalmod(mdat), xlim = c(0, 0.25))

knitr::kable(pauc(mpartial))
```

## Next

- [AUC and other curve summaries](metrics-auc.html)
- [Customizing plots](plots-customizing.html)
