---
title: "Plots overview"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

Every `precrec` object knows how to draw itself two ways.

```{r}
library(precrec)
library(ggplot2)

curves <- evalmod(scores = P10N10$scores, labels = P10N10$labels)
```

## plot or autoplot

| Function | Draws with | Use it when |
|---|---|---|
| `plot()` | base graphics | You want a picture with no dependencies |
| `autoplot()` | `ggplot2` | You want to restyle, relabel or combine the result |

```{r}
plot(curves)
```

`autoplot()` returns a `ggplot` object, so anything you can do to a plot from
`ggplot2` you can do to this one.

```{r}
autoplot(curves)
```

## Choosing what to draw

The second argument picks the curve or the metric.

```{r}
autoplot(curves, "PRC")
```

For `mode = "basic"` objects it names metrics instead, one panel each. See
[basic metric plots](plots-basic-metrics.html).

## Arguments both functions share

| Argument | Effect |
|---|---|
| `type` | `"l"` lines, `"p"` points, `"b"` both |
| `show_cb` | Draw the confidence band around an average curve |
| `raw_curves` | Draw every test set instead of the average |
| `show_legend` | Show or hide the legend |
| `reduce_points` | Thin the points before drawing; on by default |

```{r}
autoplot(curves, "PRC", type = "b")
```

## The four pages

- [ROC and precision-recall curves](plots-roc-prc.html) - the main plot
- [Basic metric plots](plots-basic-metrics.html) - metrics against
  normalized rank
- [One metric against another](plots-metric-curve.html) - free-form pairs
- [Confidence bands](plots-confidence-bands.html) and [partial
  curves](plots-partial-curves.html) - the two variations

And [customizing plots](plots-customizing.html) for changing what comes out.
