---
title: "Basic measure plots"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4
)
```

`mode = "basic"` plots evaluation measures against the normalized rank of the
scores - that is, against how far down the ranked list the cutoff sits.

```{r}
library(precrec)
library(ggplot2)

points <- evalmod(scores = P10N10$scores, labels = P10N10$labels,
  mode = "basic"
)
```

## Pick the panels

Name the measures you want. Each becomes a panel.

```{r}
autoplot(points, c("sensitivity", "specificity", "precision"))
```

```{r}
autoplot(points, c("mcc", "fscore"))
```

Called with no measures, you get all fourteen default panels at once, which
is useful for a first look and too dense for a report.

## Reading the x axis

The x axis runs from 0 to 1 and is the fraction of the dataset above the
cutoff. At x = 0 nothing is predicted positive; at x = 1 everything is. So
the left edge is the strictest cutoff and the right edge the most permissive.

This is what makes the panels comparable across datasets of different sizes.

## Scores and labels

Two extra panels show the data behind the measures rather than a measure: the
score at each rank, and the observed label.

```{r}
autoplot(points, c("score", "label"))
```

The label panel is the quickest way to see whether the positives really are
concentrated at the top of the ranking.

## Extra measures

Anything beyond the default fourteen is requested with `metrics =` and then
plotted the same way.

```{r}
extra <- evalmod(scores = P10N10$scores, labels = P10N10$labels,
  mode = "basic", metrics = c("fpr", "lift")
)

autoplot(extra, c("fpr", "lift"))
```

Asking to plot a measure that was not calculated is an error that names the
argument to add. See the [measures overview](measures-overview.html).

## Axis ranges

Measures that can go negative - `mcc`, `kappa`, `informedness`, `markedness`
and `label` - are drawn on a -1 to 1 axis. The rest use 0 to 1, and the few
with no natural bound (`lift`, `odds`, `chisq`, `cost`, `score`) are scaled
to the data.

## Next

- [One measure against another](plots-metric-curve.html)
- [Measures overview](measures-overview.html)
